### 2017-06-05 abstract.tex
empty draft of paper

\documentclass[
    ,final            % use final for the camera ready runs
%%  ,draft            % use draft while you are working on the paper
%%  ,numberedheadings % uncomment this option for numbered sections
%%  ,                 % add further options here if necessary
    ,sort&compress
  ]
  {aipproc}

\layoutstyle{8x11double}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{color}
\usepackage{amsmath}
\usepackage{units}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% FRONTMATTER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{Paradigms in Physics 2.0}

\classification{} % <Replace this text with PACS numbers; choose from this list:
                  % \texttt{http://www.aip.org/pacs/index.html}>
\keywords{curriculum redesign} % <Enter Keywords here>

\author{David Roundy}{
  address={Department of Physics, Oregon State
    University, Corvallis, Oregon 97331}}
\author{Liz Gire}{
  address={Department of Physics, Oregon State
    University, Corvallis, Oregon 97331}}
\author{Ethan Minot}{
  address={Department of Physics, Oregon State
    University, Corvallis, Oregon 97331}}
\author{Emily van Zee}{
  address={Department of Physics, Oregon State
    University, Corvallis, Oregon 97331}}
\author{Tevian Dray}{
  address={Department of Mathematics, Oregon State
    University, Corvallis, Oregon 97331}}
\author{Corinne A. Manogue}{
  address={Department of Physics, Oregon State
    University, Corvallis, Oregon 97331}}

\newcommand\myderiv[3]{\ensuremath{\left(\dfrac{\partial #1}{\partial #2}\right)_{#3}}}
\newcommand\myderivinline[3]{\ensuremath{(\slfrac{\partial #1}{\partial #2})_{#3}}}
\newcommand\inexactd{d\hspace{-0.9ex}^-\!}

\newcommand\todo[1]{\textcolor{red}{(todo: #1)}}
\newcommand\fixme[1]{\textcolor{red}{(fixme: #1)}}
\newcommand\wording[2][red]{\textcolor{#1}{\underline{#2}}}
\newcommand\wordingnote[3][red]{\textcolor{#1}{\underline{#2}\footnote{#3}}}

\definecolor{small}{rgb}{0, 0, 1}

\newcommand\dbar{{\mkern6mu\mathchar'26\mkern-12mu d}}

\begin{abstract}
In 2016, the Department of Physics at Oregon State University began a
process to revise our Paradigms in Physics curriculum for physics
majors.  This poster will describe both the process by which our
department achieved consensus on this curricular change, and the
resulting curriculum.  The Paradigms 2.0 committee begain with a
survey of students and faculty, followed by individual interviews with
the faculty teaching each existing course.  As we developed a plan to
address student- and faculty-identified challenges in the curriculum,
we met with each faculty member individually to explain and refine our
proposal, which was unanimously approved by the faculty.  Major
changes include elimination or major changes to several courses (math
methods, computational physics, modern physics, electronics, and
classical mechanics), including the introduction of two sophomore-year
courses designed specifically to help prepare students to for their
upper-division courses.
\end{abstract}

\maketitle

%% \section{Introduction}\label{intro}

%% \begin{theacknowledgments}
%%   %% NSF DUE 0618877, 0837829, 1023120, and 1323800.
%%   The funding for this project was provided, in part, by the National
%%   Science Foundation under Grant Nos.  DUE 0618877, DUE 0837829,
%%   DUE 1023120, and DUE 1323800.
%% \end{theacknowledgments}

%% \bibliographystyle{aipproc}   % if natbib is available
%% \bibliography{paper}% Produces the bibliography via BibTeX.
% \input{paper.bbl}

\end{document}

### 2017-06-05 paper.tex
empty draft of paper

% Template created by Adrienne Traxler (adrienne.traxler@wright.edu)
% Last modified: 3/22/17: Add clarifying note about why the template has too much space around its section headers.
% 
% Changelog
% 3/14/17: add 2-column figure example, tweak text to align with Word sample file, add hyperref for links/cross-references
% 6/21/16: 	Fix old PACS option and superscript affiliations.
% 5/5/16: 	Update sample table to use the REVTEX ruledtabular environment for auto-sizing;
%   		switch to using pra option until the prstper style is fixed (who knows when).
% 5/10/16: Swap suggested department/institution ordering in \affiliation lines.
% 5/12/16: Remove PACS line (deprecated).
% 6/21/16:	Switch to noshowpacs and add superscriptaddress in documentclass options.
% 
% This is my best effort to follow the Physical Review style guide plus specific changes 
% required for PERC (mostly, omitting article titles in references). This version compiles 
% with pdfLaTeX alone; using a proper .bib file changes the bibliography part at the end 
% and would require running BibTex as well.
%
% Finally, there's extra padding around section headers if you compile the bare template. 
% I believe that's because LaTeX stretches its white space to keep the floats (tables and 
% figures) near their input locations. The excess spacing goes away when a normal amount 
% of body text is filled in.

% Big list of reference documentation is here: http://journals.aps.org/revtex, 
%  see especially the APS Author Guide PDF link.

% Notes on the revtex4-1 documentclass options used: 
%  reprint does the double-column, publication ready appearance
%  prstper style formats reference numbers incorrectly, fast fix is to use pra instead.
%  Add longbibliography option to show article titles (and see note in bibliography)

%\documentclass[english,aps,prstper,reprint,showpacs]{revtex4-1}   % This version should work (prstper option), but actually formats reference numbers incorrectly.
\documentclass[english,aps,pra,reprint,noshowpacs,superscriptaddress]{revtex4-1}   
% Using pra instead of prstper gives correct square-bracket (not superscript) reference formatting
\usepackage[T1]{fontenc}	% should generally be included for better accented-word behavior
\usepackage[latin9]{inputenc}	% should generally be included for better accent behavior
\usepackage{geometry}		% for controlling page margins
\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=0.75in,rmargin=0.75in}	% define margins
\usepackage{graphicx}
\usepackage[above,below]{placeins}	% allows use of \FloatBar­rier command to force section barriers
\usepackage{times}
% Next three lines are optional, use the hyperref package to make URLs and reference links live.
\usepackage{hyperref}  
\hypersetup{colorlinks=true,urlcolor=blue,citecolor=blue,linkcolor=blue}   
\urlstyle{same}
\pagestyle{empty}			% page numbers added later, when compiling the whole proceedings
\begin{document}

\title{Paradigms in Physics 2.0}
\author{David Roundy}
\author{Liz Gire}
\author{Ethan Minot}
\author{Emily van Zee}
\affiliation{Department of Physics, Oregon State University, Corvallis, Oregon, 97331}
\author{Tevian Dray}
\affiliation{Department of Mathematics, Oregon State University, Corvallis, Oregon, 97331}
\author{Corinne A. Manogue}
\affiliation{Department of Physics, Oregon State University, Corvallis, Oregon, 97331}

%\keywords{}

\begin{abstract}
In 2016, the Department of Physics at Oregon State University began a
process to revise our Paradigms in Physics curriculum for physics
majors.  This poster will describe both the process by which our
department achieved consensus on this curricular change, and the
resulting curriculum.  The Paradigms 2.0 committee begain with a
survey of students and faculty, followed by individual interviews with
the faculty teaching each existing course.  As we developed a plan to
address student- and faculty-identified challenges in the curriculum,
we met with each faculty member individually to explain and refine our
proposal, which was unanimously approved by the faculty.  Major
changes include elimination or major changes to several courses (math
methods, computational physics, modern physics, electronics, and
classical mechanics), including the introduction of two sophomore-year
courses designed specifically to help prepare students to for their
upper-division courses.
\end{abstract}

\maketitle

\section{Introduction}
The introduction should clearly explain the nature and scope of the research as well as provide the reader with a roadmap of the rest of the paper.

Times New Roman 10 point font is used throughout the paper with single line spacing and the text is fully justified.  Paragraph style: indent first line three spaces (0.19''); there is no extra line spacing between paragraphs.  This is the standard font and layout for the individual paragraphs. % REVTeX with the options above handles this automatically.

The paper size is 8.5 $\times$ 11 inches.  Margins are 1'' top and bottom, 0.75'' left and right.  Each column is 3.35'' with 0.3'' spacing between them.

Please do not adjust the spacing of the headers.  They should have a 12pt space before and a 6pt space after each heading. (Note: \LaTeX sets the white space itself, and sometimes pads section headers. This is most likely to happen in documents that are underfilled with text---like this template---and most of those problems resolve themselves as the paper reaches a normal length.)

Information about the style of paper and answers to other submission questions can be found on the PER--Central website~\cite{per-cent}. % Official style for references is to put them inside any following punctuation, with a space between the word and the bracketed reference number. Using a ~ as in the example here inserts a non-line-breaking space.

\section{Tables and Figures}

Figures, tables, and equations must be inserted in the text and may not be grouped at the end of the paper. Important: miscounting of figures, tables, or equations may result from revisions. Please double check the numbering of these elements before you submit your paper to the volume editor.  % LaTeX handles this automatically via the \label and \ref commands.

\subsection{Figures}

If your manuscript contains figures, they are typically placed in one column (see Fig.\ \ref{fig1}).  A large figure might span two columns (see Fig.\ \ref{fig2}).  Cite all figures in the text consecutively. The word ``Figure'' should be spelled out if it is the first word of the sentence and abbreviated as ``Fig.'' elsewhere in the text. Place the figures as close as possible to their first mention in the text at the top or bottom of the page.

\begin{figure}
%\includegraphics[width=0.8\linewidth]{figure1}
\caption{Figure captions appear below the figure while table captions appear above the table.\label{fig1}}
\end{figure}

\subsubsection{Color figures}
Beginning in 2013, color figures are allowable since the PERC Proceedings will be available only in an online format.  However, authors are encouraged to check that the figures print in black and white without loss of clarity.

% Note: Revtex puts two-column figures at the top of a page. If you really need 
% to change this, gird yourself for battle and see http://tinyurl.com/j2wnoxb.
\begin{figure*}
%\includegraphics{figure2}
\caption{A 2-column figure. Center figure captions if they run one line only, and justify captions if they are multi-line.\label{fig2}}
\end{figure*}

\subsection{Tables}
Give each table a concise title and define any abbreviations in footnotes to the table. Place tables at the top or bottom of the page, close to where they are first cited (like Table \ref{tab1}). Center the table in the width of the text area, and center the caption as well.

Tables should not break across the page or column, but rather should be together cohesively.  In some cases, that may mean that a small amount of white space appears near the bottom of a column.

% REVTeX provides a ruledtabular environment that automatically sizes tables, adds double 
% lines, and adjusts column spacing. Use it around the regular tabular environment.
% For a two-column table, use \begin{table*} instead (with matching \end{table*}.
\begin{table}[htbp]
\caption{This is a sample table.\label{tab1}}
\begin{ruledtabular}
\begin{tabular}{cl}
 & \textbf{MS Word} \\ 
 & \textbf{US Letter Size Paper} \\ 
 \hline
Primary Heading & Roman numerals, all capital letters \\
First subheading & Capital letters, only first \\
 & letter of first word capitalized \\
Second subheading & Numbers, italicized 
\end{tabular}
\end{ruledtabular}
\end{table}

% Formatting tweak if needed--FloatBarrier forces floats to show here, before next section
%\FloatBarrier

\section{Other issues}

\subsection{Copyright transfer}
All accepted papers will be made available under a Creative Commons Attribution 3.0 License (CC-BY).

\subsection{Citations}
When you are preparing references, please use the PRST--PER Style Guide~\cite{style-guide}.  This document contains information about how references should be formatted, including proper journal abbreviations.  Examples directly from that style guide of a journal, book, proceedings paper, and dissertation are also provided here~\cite{smith-brown,smith,smith-proc,smith-diss}.

\subsection{Permissions}
To use previously published material from a book or journal, you must obtain written permission from the owner of the rights to the material (the original publisher and/or author). It is your responsibility to obtain permission to use copyrighted material. The executed permissions need to be sent along with the manuscript to your volume editor. Most publishers offer submission of permission requests online or via email, which may be the fastest and most convenient way of receiving a reply. Some examples with relevant links are:

\begin{itemize}
\item \url{http://www.elsevier.com/locate/permissions}
\item \url{http://www.ieee.org}
\item \url{http://www.nature.com}
\item \url{http://www.sciencemag.org}
\end{itemize}

You may also use the Permission Request Form to request permission to reprint text, tables or figures. You may complete this form and fax it to the publisher or author of the material you wish to use. A blank form is available for download click on Forms. When the signed permission is returned to you, please insert any necessary credit lines in your figure or table legends.

\section{Conclusions}
This template was newly updated for the 2016 PERC Proceedings.  The
editors apologize if any errors exist, and encourage you to contact
them with changes and other suggestions.  \cite{manogue2001paradigms}

\acknowledgments{Put references below the acknowledgements (and appendixes, if any).}

%% % For short, simple bibliographies, manually formatting works:
%% % Remember that you'll need to run pdfLaTeX twice to get the references to show (first 
%% % pass will insert ?? in their places).
%% \begin{thebibliography}{99}  
%% % number in curly brackets sets the widest possible list numbering; 99 is commonly used
%% \bibitem{per-cent} \url{http://www.per-central.org/perc/AuthorSubmission.cfm}. Retrieved 6/20/2015.

%% \bibitem{style-guide} \url{http://journals.aps.org/files/styleguide-pr.pdf}. Retrieved 5/5/2016.

%% \bibitem{smith-brown}  J.M. Smith and R. Brown, Phys. Rev. B {\bf 26}, 1 (1982)

%% \bibitem{smith} J.M. Smith, {\em Molecular Dynamics} (Academic, New York, 1980), Vol. 2, p. 20.

%% \bibitem{smith-proc} J.M. Smith, in {\em Proceedings of the International Conference on Low Temperature Physics, Madison, 1958},  edited by C. Brown (University of Wisconsin, Madison, 1958), p. 201.

%% \bibitem{smith-diss} J.M. Smith, Ph.D. thesis, Brown University, 1980.
%% % Also, maybe J. M. Smith should branch out and cite someone else for a change.
%% \end{thebibliography}

% For a longer bibliography, delete the thebibliography block above, then comment in 
% these two lines to use a .bib file with BibTeX.
\bibliographystyle{apsrev}  	% supercedes the longbibliography option, so leave commented out if you want to display article titles
\bibliography{paper}  	% don't include the .bib suffix

\end{document}

### 2017-06-06 paper.tex
cut comments citations
various minor tweaks and add sections

@@
 \section{Introduction}
-The introduction should clearly explain the nature and scope of the research as well as provide the reader with a roadmap of the rest of the paper.
+The Paradigms in Physics program
 
-Times New Roman 10 point font is used throughout the paper with single line spacing and the text is fully justified.
-Paragraph style: indent first line three spaces (0.19''); there is no extra line spacing between paragraphs.
-This is the standard font and layout for the individual paragraphs.
-% REVTeX with the options above handles this automatically.
+\subsection{Background}
+\emph{What is the Paradigms now?}
+\subsection{History of the Paradigms}
+\emph{How did we get here?~\cite{manogue2001paradigms}}
+\subsection{Motivation for the change}
+\emph{Why change? Why now? Should you do the same?}
 
-The paper size is 8.5 $\times$ 11 inches.  Margins are 1'' top and bottom, 0.75'' left and right.  Each column is 3.35'' with 0.3'' spacing between them.
+\section{Paradigms 2.0 process}
+\emph{How did we do it?}
 
-Please do not adjust the spacing of the headers.
-They should have a 12pt space before and a 6pt space after each heading.
-(Note: \LaTeX sets the white space itself, and sometimes pads section headers.
-This is most likely to happen in documents that are underfilled with text---like this template---and most of those problems resolve themselves as the paper reaches a normal length.)
+\section{Changes made}
+\emph{What did we change?}
 
-Information about the style of paper and answers to other submission questions can be found on the PER--Central website~\cite{per-cent}.
-% Official style for references is to put them inside any following punctuation, with a space between the word and the bracketed reference number.
-Using a ~ as in the example here inserts a non-line-breaking space.
+\subsection{Math bits}
+\emph{What to do with math methods?}
+\subsection{Electronics}
+\emph{Reduce quantity of electronics required, clarify expectations.}
+\subsection{5-week Paradigms}
+\emph{Why lengthen it?}
+\subsection{Sophomore courses}
+\emph{Give students more preparation when possible.  Replace Modern
+  Physics and Classical Capstone courses.}
+\subsection{Reordering}
+\emph{Improve interaction with requirements.}
+
+\section{Outcomes}
+\emph{How has it gone?}
+
+\subsection{Faculty vote}
+\emph{Unanimous result.}
+\subsection{Transition year}
+\emph{Gradual transition, getting courses taught.}
+
+\section{Future research}
+\emph{Document learning trajectories, new courses.}
 
@@
 
-Figures, tables, and equations must be inserted in the text and may not be grouped at the end of the paper.
-Important: miscounting of figures, tables, or equations may result from revisions.
-Please double check the numbering of these elements before you submit your paper to the volume editor.
+Figures, tables, and equations must be inserted in the text and may
+not be grouped at the end of the paper. Important: miscounting of
+figures, tables, or equations may result from revisions. Please double
+check the numbering of these elements before you submit your paper to
+the volume editor.
 % LaTeX handles this automatically via the \label and \ref commands.
@@
 
-If your manuscript contains figures, they are typically placed in one column (see Fig.\ \ref{fig1}).
-A large figure might span two columns (see Fig.\ \ref{fig2}).
-Cite all figures in the text consecutively.
-The word ``Figure'' should be spelled out if it is the first word of the sentence and abbreviated as ``Fig.'' elsewhere in the text.
-Place the figures as close as possible to their first mention in the text at the top or bottom of the page.
+If your manuscript contains figures, they are typically placed in one
+column (see Fig.\ \ref{fig1}).  A large figure might span two columns
+(see Fig.\ \ref{fig2}).  Cite all figures in the text
+consecutively. The word ``Figure'' should be spelled out if it is the
+first word of the sentence and abbreviated as ``Fig.'' elsewhere in
+the text. Place the figures as close as possible to their first
+mention in the text at the top or bottom of the page.
 
@@
 \subsubsection{Color figures}
-Beginning in 2013, color figures are allowable since the PERC Proceedings will be available only in an online format.
-However, authors are encouraged to check that the figures print in black and white without loss of clarity.
+Beginning in 2013, color figures are allowable since the PERC
+Proceedings will be available only in an online format.  However,
+authors are encouraged to check that the figures print in black and
+white without loss of clarity.
 
@@
   %\includegraphics{figure2}
-\caption{A 2-column figure. Center figure captions if they run one line only, and justify captions if they are multi-line.\label{fig2}}
+  picture here of the new curriculum
+\caption{A 2-column figure. Center figure captions if they run one
+  line only, and justify captions if they are multi-line.\label{fig2}}
 \end{figure*}
-
-\subsection{Tables}
-Give each table a concise title and define any abbreviations in footnotes to the table.
-Place tables at the top or bottom of the page, close to where they are first cited (like Table \ref{tab1}).
-Center the table in the width of the text area, and center the caption as well.
-
-Tables should not break across the page or column, but rather should be together cohesively.  In some cases, that may mean that a small amount of white space appears near the bottom of a column.
 
@@
 \subsection{Citations}
-When you are preparing references, please use the PRST--PER Style Guide~\cite{style-guide}.
-This document contains information about how references should be formatted, including proper journal abbreviations.
-Examples directly from that style guide of a journal, book, proceedings paper, and dissertation are also provided here~\cite{smith-brown,smith,smith-proc,smith-diss}.
+When you are preparing references, please use the PRST--PER Style
+Guide~\cite{style-guide}.  This document contains information about
+how references should be formatted, including proper journal
+abbreviations.  Examples directly from that style guide of a journal,
+book, proceedings paper, and dissertation are also provided
+here~\cite{smith-brown,smith,smith-proc,smith-diss}.
 
 \subsection{Permissions}
-To use previously published material from a book or journal, you must obtain written permission from the owner of the rights to the material (the original publisher and/or author).
-It is your responsibility to obtain permission to use copyrighted material.
-The executed permissions need to be sent along with the manuscript to your volume editor.
-Most publishers offer submission of permission requests online or via email, which may be the fastest and most convenient way of receiving a reply.
-Some examples with relevant links are:
+To use previously published material from a book or journal, you must
+obtain written permission from the owner of the rights to the material
+(the original publisher and/or author). It is your responsibility to
+obtain permission to use copyrighted material. The executed
+permissions need to be sent along with the manuscript to your volume
+editor. Most publishers offer submission of permission requests online
+or via email, which may be the fastest and most convenient way of
+receiving a reply. Some examples with relevant links are:
 
@@
 
-You may also use the Permission Request Form to request permission to reprint text, tables or figures.
-You may complete this form and fax it to the publisher or author of the material you wish to use.
-A blank form is available for download click on Forms.
-When the signed permission is returned to you, please insert any necessary credit lines in your figure or table legends.
+You may also use the Permission Request Form to request permission to
+reprint text, tables or figures. You may complete this form and fax it
+to the publisher or author of the material you wish to use. A blank
+form is available for download click on Forms. When the signed
+permission is returned to you, please insert any necessary credit
+lines in your figure or table legends.
 
@@
 editors apologize if any errors exist, and encourage you to contact
-them with changes and other suggestions.  \cite{manogue2001paradigms}
+them with changes and other suggestions.
 
 \acknowledgments{Put references below the acknowledgements (and appendixes, if any).}
-
-%% % For short, simple bibliographies, manually formatting works:
-%% % Remember that you'll need to run pdfLaTeX twice to get the references to show (first 
-%% % pass will insert ?? in their places).
-%% \begin{thebibliography}{99}  
-%% % number in curly brackets sets the widest possible list numbering; 99 is commonly used
-%% \bibitem{per-cent} \url{http://www.per-central.org/perc/AuthorSubmission.cfm}. Retrieved 6/20/2015.
-
-%% \bibitem{style-guide} \url{http://journals.aps.org/files/styleguide-pr.pdf}. Retrieved 5/5/2016.
-
-%% \bibitem{smith-brown}  J.M. Smith and R. Brown, Phys. Rev. B {\bf 26}, 1 (1982)
-
-%% \bibitem{smith} J.M. Smith, {\em Molecular Dynamics} (Academic, New York, 1980), Vol. 2, p. 20.
-
-%% \bibitem{smith-proc} J.M.
-Smith, in {\em Proceedings of the International Conference on Low Temperature Physics, Madison, 1958},  edited by C.
-Brown (University of Wisconsin, Madison, 1958), p.
-201.
-
-%% \bibitem{smith-diss} J.M. Smith, Ph.D. thesis, Brown University, 1980.
-%% % Also, maybe J. M. Smith should branch out and cite someone else for a change.
-%% \end{thebibliography}
 

### 2017-06-06 paper.tex
add tables for new schedule

@@
 
+%\newcommand\mathcourse[2]{\emph{#1 (#2)}}
+\newcommand\mathcourse[2]{\emph{#1}}
+\newcommand\noted[2]{\textbf{#1} (#2)}
+\newcommand\paradigm[1]{#1 (3)}
+\newcommand\capstone[1]{#1 (3)}
+\newcommand\onecredit[1]{#1 (1)}
+\newcommand\threecredit[1]{#1 (3)}
+\newcommand\fourcredit[1]{#1 (4)}
+
+\begin{table*}[htbp]
+\caption{Hypothetical student schedule for a student entering without
+  calculus.  Students with AP Calculus can take the junior-year
+  courses one year earlier, and students who start calculus-based
+  physics in the Fall of their second year are not
+  delayed.\label{schedule}}
+\begin{ruledtabular}
+\begin{tabular}{rlll}
+  \textbf{Year} & \textbf{Fall} & \textbf{Winter} & \textbf{Spring} \\
+  \hline
+  1 & \mathcourse{Differential calculus}{251} &
+  \mathcourse{Integral calculus}{252} &
+  \mathcourse{Multivariable calculus}{254} \\
+    & & \onecredit{Freshman Seminar} & \fourcredit{Calc Physics I} \\
+\hline  2 & \mathcourse{Vector calculus}{255}
+  & \mathcourse{Differential equations}{256}
+  & \mathcourse{Linear algebra}{341} \\
+  & \fourcredit{Calc Physics II} & \fourcredit{Calc Physics III} &\mathcourse{Series and
+    sequences}{253} \\
+  && \noted{Challenges}{3} & \noted{Theoretical Mechanics}{3}
+  \\
+\hline 3 & \paradigm{Quantum fundamentals} & \paradigm{Osciallations and waves} & \paradigm{Static fields}
+\\
+  & \paradigm{Energy and entropy} & \paradigm{Periodic systems} & \paradigm{Central forces}
+\\
+  & \onecredit{Computational lab I} & \onecredit{Computational lab II} & \onecredit{Computational lab III}
+\\
+  & \threecredit{Electronics} & \threecredit{Computer interfacing} & \onecredit{Research}
+\\
+\hline 4 & \capstone{Electromagnetism capstone} & \capstone{Quantum capstone} & \capstone{Thermal capstone}
+\\
+& \onecredit{Thesis} & \onecredit{Thesis} & \onecredit{Thesis}
+\\
+& \onecredit{Research} & \onecredit{Research}
+\end{tabular}
+\end{ruledtabular}
+\end{table*}
+
+\begin{table*}[htbp]
+\caption{Hypothetical transfer student schedule.\label{schedule}}
+\begin{ruledtabular}
+\begin{tabular}{rlll}
+  \textbf{Year} & \textbf{Fall} & \textbf{Winter} & \textbf{Spring} \\
+  \hline
+3 & \paradigm{Quantum fundamentals} & \paradigm{Osciallations and waves} & \paradigm{Static fields}
+\\
+  & \paradigm{Energy and entropy} & \paradigm{Periodic systems} & \paradigm{Central forces}
+\\
+  & \onecredit{Computational lab I} & \onecredit{Computational lab II} & \onecredit{Computational lab III}
+\\
+  & \threecredit{Electronics} & \threecredit{Computer interfacing} & \onecredit{Research}
+\\
+&& \noted{Challenges}{3} & \noted{Theoretical Mechanics}{3}
+\\
+\hline 4 & \capstone{Electromagnetism capstone} & \capstone{Quantum capstone} & \capstone{Thermal capstone}
+\\
+& \onecredit{Thesis} & \onecredit{Thesis} & \onecredit{Thesis}
+\\
+& \onecredit{Research} & \onecredit{Research}
+\end{tabular}
+\end{ruledtabular}
+\end{table*}
+
 % Formatting tweak if needed--FloatBarrier forces floats to show here, before next section

### 2017-06-07 paper.tex
work on text

@@
 \subsection{Motivation for the change}
-\emph{Why change? Why now? Should you do the same?}
+\paragraph{Curricular changes}
+There were a number of factors that motivated us to embark on the
+Paradigms~2.0 process.  The original Paradigms in Physics curricular
+reform from 200? was showing its age.  In the intervening years we had
+made a large number of modifications, such as reordering Paradigms,
+introducing one new Paradigm, and introducing and changing
+computational physics courses.  However, there were other changes
+needed that could not be separated from an examination of the
+curriculum as a whole.  Most notable were discussions of the
+requirements for computational physics, the amount of required
+electronics credits, and the role of our modern physics course.  These
+persistent issues were not easily addressed without addressing the
+physics major curriculum as a whole.
+
+\paragraph{New faculty}
+At this stage, (XX) of our faculty was not present during the orignal
+Paradigms effort.  Many of the newer faculty have only a superficial
+undertanding of our course structure.  This poses a challenge when
+these professors teach courses that are closely intertwined with one
+another.
 
@@
 
-\section{Changes made}
-\emph{What did we change?}
-
-\subsection{Math bits}
-\emph{What to do with math methods?}
-\subsection{Electronics}
-\emph{Reduce quantity of electronics required, clarify expectations.}
-\subsection{5-week Paradigms}
-\emph{Why lengthen it?}
-\subsection{Sophomore courses}
-\emph{Give students more preparation when possible.  Replace Modern
-  Physics and Classical Capstone courses.}
-\subsection{Reordering}
-\emph{Improve interaction with requirements.}
-
-\section{Outcomes}
-\emph{How has it gone?}
-
-\subsection{Faculty vote}
-\emph{Unanimous result.}
-\subsection{Transition year}
-\emph{Gradual transition, getting courses taught.}
-
-\section{Future research}
-\emph{Document learning trajectories, new courses.}
-
-\section{Tables and Figures}
-
-Figures, tables, and equations must be inserted in the text and may
-not be grouped at the end of the paper. Important: miscounting of
-figures, tables, or equations may result from revisions. Please double
-check the numbering of these elements before you submit your paper to
-the volume editor.
-% LaTeX handles this automatically via the \label and \ref commands.
-
-\subsection{Figures}
-
-If your manuscript contains figures, they are typically placed in one
-column (see Fig.\ \ref{fig1}).  A large figure might span two columns
-(see Fig.\ \ref{fig2}).  Cite all figures in the text
-consecutively. The word ``Figure'' should be spelled out if it is the
-first word of the sentence and abbreviated as ``Fig.'' elsewhere in
-the text. Place the figures as close as possible to their first
-mention in the text at the top or bottom of the page.
-
-\begin{figure}
-%\includegraphics[width=0.8\linewidth]{figure1}
-\caption{Figure captions appear below the figure while table captions appear above the table.\label{fig1}}
-\end{figure}
-
-\subsubsection{Color figures}
-Beginning in 2013, color figures are allowable since the PERC
-Proceedings will be available only in an online format.  However,
-authors are encouraged to check that the figures print in black and
-white without loss of clarity.
-
-% Note: Revtex puts two-column figures at the top of a page. If you really need 
-% to change this, gird yourself for battle and see http://tinyurl.com/j2wnoxb.
-\begin{figure*}
-  %\includegraphics{figure2}
-  picture here of the new curriculum
-\caption{A 2-column figure. Center figure captions if they run one
-  line only, and justify captions if they are multi-line.\label{fig2}}
-\end{figure*}
-
-% REVTeX provides a ruledtabular environment that automatically sizes tables, adds double 
-% lines, and adjusts column spacing. Use it around the regular tabular environment.
-% For a two-column table, use \begin{table*} instead (with matching \end{table*}.
-\begin{table}[htbp]
-\caption{This is a sample table.\label{tab1}}
-\begin{ruledtabular}
-\begin{tabular}{cl}
- & \textbf{MS Word} \\
- & \textbf{US Letter Size Paper} \\
- \hline
-Primary Heading & Roman numerals, all capital letters \\
-First subheading & Capital letters, only first \\
- & letter of first word capitalized \\
-Second subheading & Numbers, italicized 
-\end{tabular}
-\end{ruledtabular}
-\end{table}
-
 %\newcommand\mathcourse[2]{\emph{#1 (#2)}}
@@
 \newcommand\noted[2]{\textbf{#1} (#2)}
-\newcommand\paradigm[1]{#1 (3)}
+\newcommand\paradigm[1]{{\sc #1} (3)}
 \newcommand\capstone[1]{#1 (3)}
@@
 \begin{table*}[htbp]
-\caption{Hypothetical transfer student schedule.\label{schedule}}
+\caption{Hypothetical transfer student schedule.  Most transfer
+  students will have a heavier schedule in their transfer year due
+  to need to catch up on required math courses.\label{schedule}}
 \begin{ruledtabular}
@@
 \\
-&& \noted{Challenges}{3} & \noted{Theoretical Mechanics}{3}
+& \mathcourse{Linear algebra}{341} & \noted{Challenges}{3} & \noted{Theoretical Mechanics}{3}
+\\
+&&\mathcourse{Vector calculus}{255}
 \\
@@
 
+\section{Changes made}
+\emph{What did we change?}
+
+\subsection{Math bits}
+\emph{What to do with math methods?}
+\subsection{Electronics}
+\emph{Reduce quantity of electronics required, clarify expectations.}
+\subsection{5-week Paradigms}
+\emph{Why lengthen it?}
+\subsection{Sophomore courses}
+\emph{Give students more preparation when possible.  Replace Modern
+  Physics and Classical Capstone courses.}
+\subsection{Reordering}
+\emph{Improve interaction with requirements.}
+
+\section{Outcomes}
+\emph{How has it gone?}
+
+\subsection{Faculty vote}
+\emph{Unanimous result.}
+\subsection{Transition year}
+\emph{Gradual transition, getting courses taught.}
+
+\section{Future research}
+\emph{Document learning trajectories, new courses.}
+
+\begin{figure}
+%\includegraphics[width=0.8\linewidth]{figure1}
+\caption{Figure captions appear below the figure while table captions appear above the table.\label{fig1}}
+\end{figure}
+
+% Note: Revtex puts two-column figures at the top of a page. If you really need 
+% to change this, gird yourself for battle and see http://tinyurl.com/j2wnoxb.
+\begin{figure*}
+  %\includegraphics{figure2}
+  picture here of the new curriculum
+\caption{A 2-column figure. Center figure captions if they run one
+  line only, and justify captions if they are multi-line.\label{fig2}}
+\end{figure*}
+
 % Formatting tweak if needed--FloatBarrier forces floats to show here, before next section
 %\FloatBarrier
-
-\section{Other issues}
-
-\subsection{Copyright transfer}
-All accepted papers will be made available under a Creative Commons Attribution 3.0 License (CC-BY).
-
-\subsection{Citations}
-When you are preparing references, please use the PRST--PER Style
-Guide~\cite{style-guide}.  This document contains information about
-how references should be formatted, including proper journal
-abbreviations.  Examples directly from that style guide of a journal,
-book, proceedings paper, and dissertation are also provided
-here~\cite{smith-brown,smith,smith-proc,smith-diss}.
-
-\subsection{Permissions}
-To use previously published material from a book or journal, you must
-obtain written permission from the owner of the rights to the material
-(the original publisher and/or author). It is your responsibility to
-obtain permission to use copyrighted material. The executed
-permissions need to be sent along with the manuscript to your volume
-editor. Most publishers offer submission of permission requests online
-or via email, which may be the fastest and most convenient way of
-receiving a reply. Some examples with relevant links are:
-
-\begin{itemize}
-\item \url{http://www.elsevier.com/locate/permissions}
-\item \url{http://www.ieee.org}
-\item \url{http://www.nature.com}
-\item \url{http://www.sciencemag.org}
-\end{itemize}
-
-You may also use the Permission Request Form to request permission to
-reprint text, tables or figures. You may complete this form and fax it
-to the publisher or author of the material you wish to use. A blank
-form is available for download click on Forms. When the signed
-permission is returned to you, please insert any necessary credit
-lines in your figure or table legends.
 

### 2017-06-08 paper.tex
start work on pretty schedule figure

@@
   & \mathcourse{Linear algebra}{341} \\
-  & \fourcredit{Calc Physics II} & \fourcredit{Calc Physics III} &\mathcourse{Series and
+  & \fourcredit{Calc Physics II} & \fourcredit{Calc Physics III} &\mathcourse{Series \&
     sequences}{253} \\
@@
   \\
-\hline 3 & \paradigm{Quantum fundamentals} & \paradigm{Osciallations and waves} & \paradigm{Static fields}
+\hline 3 & \paradigm{Quantum fundamentals} & \paradigm{Osciallations \& waves} & \paradigm{Static fields}
 \\
-  & \paradigm{Energy and entropy} & \paradigm{Periodic systems} & \paradigm{Central forces}
+  & \paradigm{Energy \& entropy} & \paradigm{Periodic systems} & \paradigm{Central forces}
 \\
@@
   \hline
-3 & \paradigm{Quantum fundamentals} & \paradigm{Osciallations and waves} & \paradigm{Static fields}
+3 & \paradigm{Quantum fundamentals} & \paradigm{Osciallations \& waves} & \paradigm{Static fields}
 \\
-  & \paradigm{Energy and entropy} & \paradigm{Periodic systems} & \paradigm{Central forces}
+  & \paradigm{Energy \& entropy} & \paradigm{Periodic systems} & \paradigm{Central forces}
 \\
@@
 \begin{figure}
-%\includegraphics[width=0.8\linewidth]{figure1}
+\includegraphics[width=\columnwidth]{schedule}
 \caption{Figure captions appear below the figure while table captions appear above the table.\label{fig1}}

### 2017-06-09 paper.tex
tweak schedule figure
more tweaks to schedule figure
drafty conclustion

@@
 
+\begin{figure*}
+\includegraphics[width=\textwidth]{schedule}
+\caption{Hypothetical student schedule for a student entering without
+  calculus.  Students with AP Calculus can take the junior-year
+  courses one year earlier, and students who start calculus-based
+  physics in the Fall of their second year are not
+  delayed.\label{schedule}}
+\end{figure*}
+
 \begin{table*}[htbp]
@@
 
-\begin{figure}
-\includegraphics[width=\columnwidth]{schedule}
-\caption{Figure captions appear below the figure while table captions appear above the table.\label{fig1}}
-\end{figure}
-
 % Note: Revtex puts two-column figures at the top of a page. If you really need 
@@
 \section{Conclusions}
-This template was newly updated for the 2016 PERC Proceedings.  The
-editors apologize if any errors exist, and encourage you to contact
-them with changes and other suggestions.
+\emph{We made some changes, and the process went well.}
 

### 2017-06-13 paper.tex
pretty figure and edits from Corinne and Henri
start outlining new courses

@@
 \includegraphics[width=\textwidth]{schedule}
-\caption{Hypothetical student schedule for a student entering without
-  calculus.  Students with AP Calculus can take the junior-year
-  courses one year earlier, and students who start calculus-based
-  physics in the Fall of their second year are not
-  delayed.\label{schedule}}
+\caption{Hypothetical student schedules for (a) a student entering
+  without calculus and (b) a transfer student.  The Dependencies
+  between courses are noted with arrows.  The area of each course is
+  proportional to the credit value of the course, with the exception
+  of math courses, which are shown in small boxes even though they are
+  4-credit courses.\label{schedule}}
 \end{figure*}

### 2017-06-13 paper.tex
a bunch more text

--- 2017-06-09 paper.tex
@@
 \subsection{Motivation for the change}
-\paragraph{Curricular changes}
 There were a number of factors that motivated us to embark on the
 Paradigms~2.0 process.  The original Paradigms in Physics curricular
-reform from 200? was showing its age.  In the intervening years we had
-made a large number of modifications, such as reordering Paradigms,
-introducing one new Paradigm, and introducing and changing
-computational physics courses.  However, there were other changes
-needed that could not be separated from an examination of the
-curriculum as a whole.  Most notable were discussions of the
-requirements for computational physics, the amount of required
-electronics credits, and the role of our modern physics course.  These
-persistent issues were not easily addressed without addressing the
-physics major curriculum as a whole.
+reform from 200? was showing its age.  In the intervening years we
+have made a large number of modifications, such as reordering
+Paradigms, introducing one new Paradigm, and introducing and changing
+computational physics courses.
+
+\paragraph{Transfer students}
+Our curriculum is strongly affected by our desire to accomodate
+transfer students from community colleges.  This leads us to focus on
+a physics major in which all upper-division courses are taken in two
+years.  However, our current curriculum is very hard on students in
+the fall of their junior year, and especially so for transfer
+students.  So we hoped to structure the courses such that non-transfer
+students could be better prepared for their junior year, while
+transfer students could postpone a few courses so that their junior
+Fall would be no more difficult than that of any other student.
+
+\paragraph{Coupled curricular changes}
+A backlog of curricular changes have accumulated that could not be
+separated from an examination of the curriculum as a whole.  Most
+notable were discussions of the requirements for computational
+physics, the amount of required electronics credits, and the role of
+our modern physics course.  These persistent issues were not easily
+addressed without addressing the physics major curriculum as a whole,
+and any discussion of one tended to devolve into a discussion of
+everything, without progress being made.
 
@@
 \includegraphics[width=\textwidth]{schedule}
-\caption{Hypothetical student schedule for a student entering without
-  calculus.  Students with AP Calculus can take the junior-year
-  courses one year earlier, and students who start calculus-based
-  physics in the Fall of their second year are not
-  delayed.\label{schedule}}
+\caption{Hypothetical student schedules for (a) a student entering
+  without calculus and (b) a transfer student.  The Dependencies
+  between courses are noted with arrows.  The area of each course is
+  proportional to the credit value of the course, with the exception
+  of math courses, which are shown in small boxes even though they are
+  4-credit courses.  The two newly introduced courses are indicated
+  with shadows.\label{fig:schedule}}
 \end{figure*}
+
+\section{Changes made}
+As a result of this process, we have implemented a number of changes
+to our physics major curriculum.  The resulting curriculum is outlined
+in Fig.~\ref{fig:schedule}.  The changes consisted of introducing two
+new sophomore-level courses (which may be taken in the junior year) to
+better prepare our majors for the Paradigms.  We eliminated Modern
+Physics and the Classical Mechanics capstone in favor of the two new
+sophomore-level courses.  We eliminated the Math Methods course in
+favor of ``math bits'' integrated into the Paradigms themselves.
+Finally, we restructured our nine 3-week Paradigms courses into six
+5-week Paradigms.  We changed our computational physics requirement,
+and reduced the number of electronics courses.
+
+\subsection{5-week Paradigms}
+To begin with the most distinctive feature of our curriculum, we
+wanted to maintain the existing intensive 7-hour-per-week schedule in
+the junior year, which we have found effective.  However, we chose to
+change from three 3-week Paradigms per quarter to two 5-week Paradigms
+per quarter.    This change has several advantages.  It gives students
+a bit more time with each professor prior to their final exam.  This
+gives a bit more flexibility, for instance, in case of illness.
+Finally, by reducing the number of Paradigms, we allow the curriculum
+to be taught with fewer faculty, albeit at a work load per professor.
+
+We put considerable thought into the content and ordering of the new
+Paradigms.  In most cases, we think of each 5-week Paradigm as either
+one formerly 4-week Paradigm (the one that consumed the extra week
+each quarter) or two 3-week Paradigms compressed.  The major
+scheduling change is to place Static Fields (electrostatics and
+magnetostatics) in the spring quarter.  This intended to give students
+a bit more time to take Vector Calculus.  It also puts all use of
+curvilinear coordinates in the spring quarter, which should ease the
+learning of Central Forces.  The final major change was the
+elimination of special relativity from the Paradigms, which is has now
+moved into the new sophomore-level Theoretical Mechanics course.
+
+\subsection{Math bits}
+We chose to eliminate the Math Methods course in favor of formal
+``math bits'' in the Paradigms.  Our previous Math Methods Capstone
+posed a challenge regarding placement in the curriculum: it needed to
+be sufficiently late in order for students to be adequately prepared
+for advanced mathematical content, but in order to be functional in
+the curriculum it should teach content that is required by other
+physics courses, and thus should be earlier in the curriculum.  In
+response to these issues, and in recognition that the majority of our
+majors are not bound for graduate school, and have no need for
+advanced math methods, we chose to integrate the needed math content
+into the Paradigms, and allow advanced students to take our
+graduate-level Math Methods course.  In order to ensure that our
+students to receive the math instruction they require, and based on
+our previous experience with math-intensive weeks (the ``odd'' weeks)
+in the Paradigms, we chose to integrate one week of ``math bits'' into
+each Paradigm.  The math bits for the entire year are treated as a
+distinct teaching assignment, so students will be accustomed to the
+treatment of math as a distinct subject supporting the physics
+content.  This provides some relief from teaching for the Paradigms
+professors, and at the same time ensures that they do not succumb to
+the temptation to short-change the math content to the detriment of
+the students.
+
+\subsection{Electronics}
+We chose to reduce the Electronics requirement from two 3-credit
+courses to one 3-credit course.  For many physics majors, this is more
+than sufficient, and gives students a greater number of electives.  In
+addition, we removed the lecture section of this course---which had
+developed an unusually high student work load for a 3-credit
+course---in favor of in-lab instruction.
+
+A final change to Electronics is that we now will \emph{require}
+Electronics during the junior year (specifically as a prerequisite for
+Oscillations and Waves).  This is made possible by the reduction in
+Fall workload for incoming transfer students, who had previously
+usually taken Electronics as a senior.  This change has enabled us to
+articulate distinct and sequenced learning outcomes from these two
+courses, particularly in the realm of complex exponentials and Fourier
+transforms.  It is also beneficial in providing students with
+trouble-shooting skills prior to the in-class labs taught in
+Oscillations and Waves.
+
+\subsection{Computational lab}
+Over the last six years, we have been developing a 1-credit
+computational laboratory course that accompanies the Paradigms.  We
+chose to require this course, while removing a requirement for a
+lower-division 3-credit course in computational physics.  This
+lower-division course was challenging to teach, since it always had a
+mix of lower- and upper-division students, with very different skill
+levels and needs.  Transfer students now take computation alongside
+the non-transfer students, in a laboratory setting using pair
+programming to help new programmers to learn.
+
+\subsection{Sophomore courses}
+We introduced two new sophomore-level courses: \emph{Challenges in
+  Contemporary Physics}, and \emph{Theoretical Mechanics}.  Both of
+these courses are designed to ramp up student mathematical abilities
+prior to their junior year.  The Challenges course will focus on
+estimation, dimensional reasoning, and interpretation of integrals,
+while Theoretical Mechanics will teach power series approximations,
+and expose students to increased levels of mathematical
+sophistication.
+
+These courses \emph{will} have the challenge (which was removed from
+the computational course!) of teaching both juniors and sophomores
+together.  They will be taught in the winter and spring quarters, so
+as to reduce the burden on transfer students in the Fall.  They were
+explicitly constructed to \emph{not} teach any content required for
+Fall or Winter Paradigms so that they can be taken concurrently with
+the Paradigms.
+
+\subsubsection{Challenges in Contemporary Physics}
+\emph{Course overview: relevant to world, benefit for diversity and
+  recruitment, replaces Modern Physics. Topics.}
+\subsubsection{Techniques of Theoretical Mechanics}
+\emph{Course overview: special relativity, rocket motion, Lagrangian
+  mechanics.}
+
+\section{Outcomes}
+\emph{How has it gone?}
+
+\subsection{Faculty vote}
+\emph{Unanimous result.}
+\subsection{Transition year}
+\emph{Gradual transition, getting courses taught.}
+
+\section{Future research}
+\emph{Document learning trajectories, new courses.}
+
+% Formatting tweak if needed--FloatBarrier forces floats to show here, before next section
+%\FloatBarrier
+
+\section{Conclusions}
+\emph{We made some changes, and the process went well.}
+
+\acknowledgments{Put references below the acknowledgements (and appendixes, if any).}
+
+% For a longer bibliography, delete the thebibliography block above, then comment in 
+% these two lines to use a .bib file with BibTeX.
+\bibliographystyle{apsrev}  	% supercedes the longbibliography option, so leave commented out if you want to display article titles
+\bibliography{paper}  	% don't include the .bib suffix
 
@@
 
-\section{Changes made}
-\emph{What did we change?}
-
-\subsection{Math bits}
-\emph{What to do with math methods?}
-\subsection{Electronics}
-\emph{Reduce quantity of electronics required, clarify expectations.}
-\subsection{5-week Paradigms}
-\emph{Why lengthen it?}
-\subsection{Sophomore courses}
-\emph{Give students more preparation when possible.  Replace Modern
-  Physics and Classical Capstone courses.}
-\subsection{Reordering}
-\emph{Improve interaction with requirements.}
-
-\section{Outcomes}
-\emph{How has it gone?}
-
-\subsection{Faculty vote}
-\emph{Unanimous result.}
-\subsection{Transition year}
-\emph{Gradual transition, getting courses taught.}
-
-\section{Future research}
-\emph{Document learning trajectories, new courses.}
-
-% Note: Revtex puts two-column figures at the top of a page. If you really need 
-% to change this, gird yourself for battle and see http://tinyurl.com/j2wnoxb.
-\begin{figure*}
-  %\includegraphics{figure2}
-  picture here of the new curriculum
-\caption{A 2-column figure. Center figure captions if they run one
-  line only, and justify captions if they are multi-line.\label{fig2}}
-\end{figure*}
-
-% Formatting tweak if needed--FloatBarrier forces floats to show here, before next section
-%\FloatBarrier
-
-\section{Conclusions}
-\emph{We made some changes, and the process went well.}
-
-\acknowledgments{Put references below the acknowledgements (and appendixes, if any).}
-
-% For a longer bibliography, delete the thebibliography block above, then comment in 
-% these two lines to use a .bib file with BibTeX.
-\bibliographystyle{apsrev}  	% supercedes the longbibliography option, so leave commented out if you want to display article titles
-\bibliography{paper}  	% don't include the .bib suffix
-
 \end{document}

### 2017-06-21 paper.tex
add history and new courses text (the latter from the grant proposal)

@@
 \usepackage{graphicx}
-\usepackage[above,below]{placeins}	% allows use of \FloatBar­rier command to force section barriers
+\usepackage[above,below]{placeins}	% allows use of \FloatBarÂ­rier command to force section barriers
 \usepackage{times}
@@
 \section{Introduction}
-The Paradigms in Physics program
+The Paradigms in Physics program began around the time when our
+current students were born~\cite{manogue2001paradigms}.  The time has
+arrived for a change.  Over the years, we have identified a number of
+problems in our curriculum that we were unable to address without
+addressing the physics major curriculum as a whole.  In this paper, we
+share the motivation for our Paradigms 2.0 effort, the process by
+which we achieved consensus on an improved curriculum, and will give
+a summary and a few highlights of the new program.
 
 \subsection{Background}
-\emph{What is the Paradigms now?}
-\subsection{History of the Paradigms}
-\emph{How did we get here?~\cite{manogue2001paradigms}}
-\subsection{Motivation for the change}
+The Paradigms in Physics project began in 1996, when three faculty
+members at Oregon State applied to the NSF for funding to redesign the
+upper-division curriculum.  The motivation for the original Paradigms
+project was in fact similar to the motivation for Paradigms 2.0: a
+desire to soften the ``brick wall'' encountered by students upon
+reaching the junior year, within the constraint that transfer students
+must be able to graduate with two years of upper-division
+courses~\cite{manogue2001paradigms}.  The process of redesigning the
+curriculum was dominated by the creation index cards listing subject
+content, and sorting these cards into courses.  This process involved
+the entire faculty, and culminated in a unanimous agreement to adopt
+the resulting curriculum.
+
+The resulting curriculum was primarily composed of junior-year
+``Paradigms'' courses, followed by senior-year ``Capstone'' courses.
+The Paradigms were intensive 2-credit, 3-week courses which meet every
+day, for a total of seven hours per week.  Paradigms incorporate
+laboratory experiences and active engagement into the class, and
+typically have two problem sets per week.  Capstones were more
+traditional 3-credit courses which meet three hours per week for an
+entire quarter.  The Paradigms courses had integrated math content,
+followed by a Math Methods course.
+
+A coupe of decades have passed since the original Paradigms effort.
+During this time we have made a number of changes: e.g. a new
+\emph{first} Paradigm was introduced to soften the beginning of the
+junior year, the order of Paradigms was shuffled more than once, Math
+Methods was moved from the beginning of the senior year to the end of
+the junior year, and a computational laboratory course was introduced
+to accompany the Paradigms.  We maintained the tradition of meeting
+every three weeks to discuss issues relating to upper-division
+teaching.  New faculty arrived and learned to teach the new courses,
+and introduced their own approach to the courses.
+
+\subsection{Motivation for Paradigms 2.0}
 There were a number of factors that motivated us to embark on the
@@
   Contemporary Physics}, and \emph{Theoretical Mechanics}.  Both of
-these courses are designed to ramp up student mathematical abilities
-prior to their junior year.  The Challenges course will focus on
-estimation, dimensional reasoning, and interpretation of integrals,
-while Theoretical Mechanics will teach power series approximations,
-and expose students to increased levels of mathematical
-sophistication.
+these courses ramp up student mathematical abilities prior to their
+junior year.  The Challenges course will focus on estimation,
+dimensional reasoning, and interpretation of integrals, while
+Theoretical Mechanics will teach power series approximations, and
+expose students to increased levels of mathematical sophistication.
 
@@
 \subsubsection{Challenges in Contemporary Physics}
-\emph{Course overview: relevant to world, benefit for diversity and
-  recruitment, replaces Modern Physics. Topics.}
+Climate change and renewable energy are major contemporary
+challenges. Students are inspired to engage with these challenges, and
+physics will play a critical role in addressing these challenges.
+Former Energy Secretary Steve Chu recently noted that ``physics majors
+seem to be increasing,'' and pointed to millennials' interests in the
+risks of climate change, clean energy, and
+sustainability~\cite{kramer2016gathering}.  However, most majors get
+little or no interaction with such challenges in their physics
+classes.
+
+We have introduced a course that fits seamlessly into the physics
+major sequence, that focuses on the contemporary challenges of
+sustainable energy and climate change. The course will be a bridge
+between introductory physics and upper-division classes for both
+physics majors and students from related disciplines such as chemistry
+and engineering. The challenges theme will motivate the introduction
+of thermal physics and quantum mechanics as well as the development of
+physical reasoning skills. We anticipate that an emphasis on the role
+of physics in overcoming societal challenges will help us recruit and
+retain a greater number of physics majors, especially women and
+under-represented minorities.
+
+At Oregon State, as at many other universities, a course in Modern
+Physics traditionally bridges the lower and upper divisions and serves
+as an introduction to quantum mechanics. Such courses typically follow
+a historical narrative, celebrating the achievements of men like
+Planck, Einstein, Rutherford, Bohr, Bragg, Millikan, de Broglie,
+Pauli, Schroedinger, and Dirac. Indeed, it is a wonderful and exciting
+story - and inspiring for those who imagine themselves as
+reincarnations of these men. However, we suggest that this historical
+narrative does not attract a wide cross-section of participants to
+physics. Instead, we have introduced a forward-looking narrative that
+emphasizes the relevance of physics today unconstrained by the details
+of the great transition from 1905 to 1945.
+
+Each physics topic is linked to contemporary challenges. An
+understanding of electromagnetic waves is required to understand how
+light interacts with vibrational modes of CO$_2$. The equipartition
+theorem explains the heat energy stored in the ocean. Quantum
+mechanics explains the blackbody spectrum, semiconductor electronic
+structure, and solar photovoltaics. The Carnot cycle convinces us that
+a 95\%-efficient gas furnace cannot compete with an efficient heat
+pump.
+
+The new perspectives associated with this course, along with an
+emphasis on expanding students' reasoning skills, will strengthen the
+connection between lower- and upper-division physics classes. Students
+will learn the importance of approximate calculations (Fermi
+estimates) that establish reasonable bounds on the exact
+answer. Students will learn that equations help us interpret the
+physical world, i.e. equations are used for much more than mindless
+algebraic manipulations.  Students will learn how integration and
+differentiation can be applied to real data - not just the perfect
+functions that are taught in math class. Students will learn how
+graphing of relationships is key tool for physicists. The utility of
+these skills will be firmly established by application to contemporary
+challenges.
+
 \subsubsection{Techniques of Theoretical Mechanics}
-\emph{Course overview: special relativity, rocket motion, Lagrangian
-  mechanics.}
+
+To physicists, mathematics is a language for telling stories about
+conceptual physical relationships.  Our understandings of conceptual
+connections guide us when we manipulate
+equations~\cite{gire2008resources, bing2007cognitive,
+  wittmann2015mathematical}[75].  For example, when solving ordinary
+differential equations by separating variables, we choose limits of
+integration so that the lower (and upper) limits for all variables
+describe the same physical state.  We monitor algebraic manipulations
+using conceptual understandings of physics quantities. For example, we
+make sure that each term in a sum has the same physical dimensions
+(e.g. you can't add a frequency to a frequency squared). We make sure
+that expressions on both sides of an equals sign have the same
+dimensions~\cite{lenz2016dimensional}. We consider limiting cases to
+make sure that a derived expression reasonably describes the system in
+extreme states~\cite{singh2002physical}. Our geometric understandings
+guide the formulation of equations~\cite{gire2014arrows}. Sometimes,
+we do not have adequate mathematics to describe complex physical
+systems. We use conceptual understandings to make choices about
+reasonable idealizations in order to create useful mathematical
+models~\cite{bing2007cognitive}.
+
+Being fluent in this physics language involves both (1) translating a
+conceptual story into a mathematical representation and (2) discerning
+a conceptual story from a mathematical
+representation~\cite{tuminaro2007elements}. At the introductory level,
+students often see concepts and mathematics as being separate. These
+students do not expect equations to help their conceptual
+understanding~\cite{adams2006new, redish1998expectations} and have
+trouble formulating equations from a conceptual analysis. Students who
+do not have mathematical fluency for physics struggle in advanced
+physics courses~\cite{manogue2009cognitive}.
+
+We propose a course that explicitly focuses on physics majors'
+mathematical fluency and sense- making skills. Introductory courses
+begin to develop these aspects, but the primary goals of those courses
+tend to be for students to apply simple concepts (Newton's Laws,
+energy conservation, etc.)  to relatively straightforward and very
+idealized problems.
+
+This course will be taken just after the introductory physics
+sequence. It will leverage students' understandings of simple
+mechanical systems from introductory physics to develop their
+mathematical fluency in the context of less idealized, more
+sophisticated systems. Each physics topic (velocity-dependent forces,
+Lagrangian mechanics, Hamiltonian mechanics, special relativity) will
+focus on specific math methods (e.g. differential equations, calculus
+of variations, generalized coordinates, Legendre transformations,
+linear transformations, hyperbola geometry) and will explicitly
+feature strategies to make physical sense of mathematics (symbolic
+problem solving, reading equations like sentences, making
+idealization, coordinating representations, limiting case analysis,
+dimensional checking).
 
@@
 
-\begin{table*}[htbp]
-\caption{Hypothetical student schedule for a student entering without
-  calculus.  Students with AP Calculus can take the junior-year
-  courses one year earlier, and students who start calculus-based
-  physics in the Fall of their second year are not
-  delayed.\label{schedule}}
-\begin{ruledtabular}
-\begin{tabular}{rlll}
-  \textbf{Year} & \textbf{Fall} & \textbf{Winter} & \textbf{Spring} \\
-  \hline
-  1 & \mathcourse{Differential calculus}{251} &
-  \mathcourse{Integral calculus}{252} &
-  \mathcourse{Multivariable calculus}{254} \\
-    & & \onecredit{Freshman Seminar} & \fourcredit{Calc Physics I} \\
-\hline  2 & \mathcourse{Vector calculus}{255}
-  & \mathcourse{Differential equations}{256}
-  & \mathcourse{Linear algebra}{341} \\
-  & \fourcredit{Calc Physics II} & \fourcredit{Calc Physics III} &\mathcourse{Series \&
-    sequences}{253} \\
-  && \noted{Challenges}{3} & \noted{Theoretical Mechanics}{3}
-  \\
-\hline 3 & \paradigm{Quantum fundamentals} & \paradigm{Osciallations \& waves} & \paradigm{Static fields}
-\\
-  & \paradigm{Energy \& entropy} & \paradigm{Periodic systems} & \paradigm{Central forces}
-\\
-  & \onecredit{Computational lab I} & \onecredit{Computational lab II} & \onecredit{Computational lab III}
-\\
-  & \threecredit{Electronics} & \threecredit{Computer interfacing} & \onecredit{Research}
-\\
-\hline 4 & \capstone{Electromagnetism capstone} & \capstone{Quantum capstone} & \capstone{Thermal capstone}
-\\
-& \onecredit{Thesis} & \onecredit{Thesis} & \onecredit{Thesis}
-\\
-& \onecredit{Research} & \onecredit{Research}
-\end{tabular}
-\end{ruledtabular}
-\end{table*}
+%% \begin{table*}[htbp]
+%% \caption{Hypothetical student schedule for a student entering without
+%%   calculus.  Students with AP Calculus can take the junior-year
+%%   courses one year earlier, and students who start calculus-based
+%%   physics in the Fall of their second year are not
+%%   delayed.\label{schedule}}
+%% \begin{ruledtabular}
+%% \begin{tabular}{rlll}
+%%   \textbf{Year} & \textbf{Fall} & \textbf{Winter} & \textbf{Spring} \\
+%%   \hline
+%%   1 & \mathcourse{Differential calculus}{251} &
+%%   \mathcourse{Integral calculus}{252} &
+%%   \mathcourse{Multivariable calculus}{254} \\
+%%     & & \onecredit{Freshman Seminar} & \fourcredit{Calc Physics I} \\
+%% \hline  2 & \mathcourse{Vector calculus}{255}
+%%   & \mathcourse{Differential equations}{256}
+%%   & \mathcourse{Linear algebra}{341} \\
+%%   & \fourcredit{Calc Physics II} & \fourcredit{Calc Physics III} &\mathcourse{Series \&
+%%     sequences}{253} \\
+%%   && \noted{Challenges}{3} & \noted{Theoretical Mechanics}{3}
+%%   \\
+%% \hline 3 & \paradigm{Quantum fundamentals} & \paradigm{Osciallations \& waves} & \paradigm{Static fields}
+%% \\
+%%   & \paradigm{Energy \& entropy} & \paradigm{Periodic systems} & \paradigm{Central forces}
+%% \\
+%%   & \onecredit{Computational lab I} & \onecredit{Computational lab II} & \onecredit{Computational lab III}
+%% \\
+%%   & \threecredit{Electronics} & \threecredit{Computer interfacing} & \onecredit{Research}
+%% \\
+%% \hline 4 & \capstone{Electromagnetism capstone} & \capstone{Quantum capstone} & \capstone{Thermal capstone}
+%% \\
+%% & \onecredit{Thesis} & \onecredit{Thesis} & \onecredit{Thesis}
+%% \\
+%% & \onecredit{Research} & \onecredit{Research}
+%% \end{tabular}
+%% \end{ruledtabular}
+%% \end{table*}
 
-\begin{table*}[htbp]
-\caption{Hypothetical transfer student schedule.  Most transfer
-  students will have a heavier schedule in their transfer year due
-  to need to catch up on required math courses.\label{schedule}}
-\begin{ruledtabular}
-\begin{tabular}{rlll}
-  \textbf{Year} & \textbf{Fall} & \textbf{Winter} & \textbf{Spring} \\
-  \hline
-3 & \paradigm{Quantum fundamentals} & \paradigm{Osciallations \& waves} & \paradigm{Static fields}
-\\
-  & \paradigm{Energy \& entropy} & \paradigm{Periodic systems} & \paradigm{Central forces}
-\\
-  & \onecredit{Computational lab I} & \onecredit{Computational lab II} & \onecredit{Computational lab III}
-\\
-  & \threecredit{Electronics} & \threecredit{Computer interfacing} & \onecredit{Research}
-\\
-& \mathcourse{Linear algebra}{341} & \noted{Challenges}{3} & \noted{Theoretical Mechanics}{3}
-\\
-&&\mathcourse{Vector calculus}{255}
-\\
-\hline 4 & \capstone{Electromagnetism capstone} & \capstone{Quantum capstone} & \capstone{Thermal capstone}
-\\
-& \onecredit{Thesis} & \onecredit{Thesis} & \onecredit{Thesis}
-\\
-& \onecredit{Research} & \onecredit{Research}
-\end{tabular}
-\end{ruledtabular}
-\end{table*}
+%% \begin{table*}[htbp]
+%% \caption{Hypothetical transfer student schedule.  Most transfer
+%%   students will have a heavier schedule in their transfer year due
+%%   to need to catch up on required math courses.\label{schedule}}
+%% \begin{ruledtabular}
+%% \begin{tabular}{rlll}
+%%   \textbf{Year} & \textbf{Fall} & \textbf{Winter} & \textbf{Spring} \\
+%%   \hline
+%% 3 & \paradigm{Quantum fundamentals} & \paradigm{Osciallations \& waves} & \paradigm{Static fields}
+%% \\
+%%   & \paradigm{Energy \& entropy} & \paradigm{Periodic systems} & \paradigm{Central forces}
+%% \\
+%%   & \onecredit{Computational lab I} & \onecredit{Computational lab II} & \onecredit{Computational lab III}
+%% \\
+%%   & \threecredit{Electronics} & \threecredit{Computer interfacing} & \onecredit{Research}
+%% \\
+%% & \mathcourse{Linear algebra}{341} & \noted{Challenges}{3} & \noted{Theoretical Mechanics}{3}
+%% \\
+%% &&\mathcourse{Vector calculus}{255}
+%% \\
+%% \hline 4 & \capstone{Electromagnetism capstone} & \capstone{Quantum capstone} & \capstone{Thermal capstone}
+%% \\
+%% & \onecredit{Thesis} & \onecredit{Thesis} & \onecredit{Thesis}
+%% \\
+%% & \onecredit{Research} & \onecredit{Research}
+%% \end{tabular}
+%% \end{ruledtabular}
+%% \end{table*}
 

### 2017-06-21 paper.tex
articulate ideas from meeting with Corinne

@@
 \section{Introduction}
+Basic Paradigms in Physics ideas:
+\begin{enumerate}
+\item Attention to sequencing of concepts (and articulating content
+  learning goals)
+\item Faculty buy in and understanding of the sequencing and goals
+\item Active engagement
+\end{enumerate}
+\emph{This} is the Paradigms.  \emph{Explain why our 2.0 changes were
+  made, how they addressed these principles.  (i.e. sophomore-year
+  courses, math bits, giving students more electives as seniors) Also
+  articulate the process (both 1.0 and 2.0) in terms of these big
+  ideas.}
+
 The Paradigms in Physics program began around the time when our

### 2017-06-27 paper.tex
input from Corinne and Ethan

@@
 
+\emph{Faculty vote}
+
 %\newcommand\mathcourse[2]{\emph{#1 (#2)}}
@@
 \subsubsection{Challenges in Contemporary Physics}
-Climate change and renewable energy are major contemporary
-challenges. Students are inspired to engage with these challenges, and
-physics will play a critical role in addressing these challenges.
-Former Energy Secretary Steve Chu recently noted that ``physics majors
-seem to be increasing,'' and pointed to millennials' interests in the
-risks of climate change, clean energy, and
-sustainability~\cite{kramer2016gathering}.  However, most majors get
-little or no interaction with such challenges in their physics
-classes.
 
-We have introduced a course that fits seamlessly into the physics
-major sequence, that focuses on the contemporary challenges of
-sustainable energy and climate change. The course will be a bridge
-between introductory physics and upper-division classes for both
-physics majors and students from related disciplines such as chemistry
-and engineering. The challenges theme will motivate the introduction
-of thermal physics and quantum mechanics as well as the development of
-physical reasoning skills. We anticipate that an emphasis on the role
-of physics in overcoming societal challenges will help us recruit and
-retain a greater number of physics majors, especially women and
-under-represented minorities.
+In ``The Physics of Contemporary Challenges'' we prepare students to
+apply physics concepts and physical reasoning skills to sustainable
+energy issues, climate change mechanisms, space exploration and
+puzzles in fundamental physics. These ``real-world'' topics are chosen
+for either the societal need (energy and climate), and/or the human
+need to explore (space and fundamental physics). By prioritizing
+inclusion of the most engaging challenges, we aim to attract and
+retain as many potential physics majors as possible.
 
-At Oregon State, as at many other universities, a course in Modern
-Physics traditionally bridges the lower and upper divisions and serves
-as an introduction to quantum mechanics. Such courses typically follow
-a historical narrative, celebrating the achievements of men like
-Planck, Einstein, Rutherford, Bohr, Bragg, Millikan, de Broglie,
-Pauli, Schroedinger, and Dirac. Indeed, it is a wonderful and exciting
-story - and inspiring for those who imagine themselves as
-reincarnations of these men. However, we suggest that this historical
-narrative does not attract a wide cross-section of participants to
-physics. Instead, we have introduced a forward-looking narrative that
-emphasizes the relevance of physics today unconstrained by the details
-of the great transition from 1905 to 1945.
+While contemporary challenges determine the narrative flow of the
+course, physics concepts and physical reasoning skills are the main
+substance. Students are introduced to thermodynamics, statistical
+mechanics, electromagnetic radiation, quantum mechanics, and modern
+experimental physics---in each case motivated by one or more
+``contemporary challenges''. Together with new physics concepts,
+students are given new reasoning tools, such as the equipartition
+theorem, the quantum-classical correspondence principle, order of
+magnitude calculations, simplifying assumptions, and numerical
+integration. Mathematically rigorous derivations are only briefly
+discussed in class, with additional details offered in online video
+format. In summary, we emphasis the physics concepts and reasoning
+skills that allow professional physicists to quickly/quantitatively
+make an initial assessment of a complex problem.
 
-Each physics topic is linked to contemporary challenges. An
-understanding of electromagnetic waves is required to understand how
-light interacts with vibrational modes of CO$_2$. The equipartition
-theorem explains the heat energy stored in the ocean. Quantum
-mechanics explains the blackbody spectrum, semiconductor electronic
-structure, and solar photovoltaics. The Carnot cycle convinces us that
-a 95\%-efficient gas furnace cannot compete with an efficient heat
-pump.
+The students engage with the material by solving daily quiz questions
+in class, solving weekly homework problems, performing a modern
+physics experiment, and writing a term paper. The daily quiz questions
+help clarify the goal of each lecture. The weekly homework questions
+are directly related to contemporary challenges are require a variety
+of estimation techniques. The modern physics experiment, performed
+with a low cost kit (LEDs, diffraction grating, rulers, volt meter
+etc.) emphasizes a quantitative approach to observation and explores
+issues of experimental uncertainty. Lastly, the term paper allows
+students to dive deeper into a particular contemporary challenge. The
+term paper tests student's ability to make their own initial
+assessment of a complex problem, and communicate their ideas with
+diagrams, graphs and estimations.
 
-The new perspectives associated with this course, along with an
-emphasis on expanding students' reasoning skills, will strengthen the
-connection between lower- and upper-division physics classes. Students
-will learn the importance of approximate calculations (Fermi
-estimates) that establish reasonable bounds on the exact
-answer. Students will learn that equations help us interpret the
-physical world, i.e. equations are used for much more than mindless
-algebraic manipulations.  Students will learn how integration and
-differentiation can be applied to real data - not just the perfect
-functions that are taught in math class. Students will learn how
-graphing of relationships is key tool for physicists. The utility of
-these skills will be firmly established by application to contemporary
-challenges.
+%% Climate change and renewable energy are major contemporary
+%% challenges. Students are inspired to engage with these challenges, and
+%% physics will play a critical role in addressing these challenges.
+%% Former Energy Secretary Steve Chu recently noted that ``physics majors
+%% seem to be increasing,'' and pointed to millennials' interests in the
+%% risks of climate change, clean energy, and
+%% sustainability~\cite{kramer2016gathering}.  However, most majors get
+%% little or no interaction with such challenges in their physics
+%% classes.
+
+%% We have introduced a course that fits seamlessly into the physics
+%% major sequence, that focuses on the contemporary challenges of
+%% sustainable energy and climate change. The course will be a bridge
+%% between introductory physics and upper-division classes for both
+%% physics majors and students from related disciplines such as chemistry
+%% and engineering. The challenges theme will motivate the introduction
+%% of thermal physics and quantum mechanics as well as the development of
+%% physical reasoning skills. We anticipate that an emphasis on the role
+%% of physics in overcoming societal challenges will help us recruit and
+%% retain a greater number of physics majors, especially women and
+%% under-represented minorities.
+
+%% At Oregon State, as at many other universities, a course in Modern
+%% Physics traditionally bridges the lower and upper divisions and serves
+%% as an introduction to quantum mechanics. Such courses typically follow
+%% a historical narrative, celebrating the achievements of men like
+%% Planck, Einstein, Rutherford, Bohr, Bragg, Millikan, de Broglie,
+%% Pauli, Schroedinger, and Dirac. Indeed, it is a wonderful and exciting
+%% story - and inspiring for those who imagine themselves as
+%% reincarnations of these men. However, we suggest that this historical
+%% narrative does not attract a wide cross-section of participants to
+%% physics. Instead, we have introduced a forward-looking narrative that
+%% emphasizes the relevance of physics today unconstrained by the details
+%% of the great transition from 1905 to 1945.
+
+%% Each physics topic is linked to contemporary challenges. An
+%% understanding of electromagnetic waves is required to understand how
+%% light interacts with vibrational modes of CO$_2$. The equipartition
+%% theorem explains the heat energy stored in the ocean. Quantum
+%% mechanics explains the blackbody spectrum, semiconductor electronic
+%% structure, and solar photovoltaics. The Carnot cycle convinces us that
+%% a 95\%-efficient gas furnace cannot compete with an efficient heat
+%% pump.
+
+%% The new perspectives associated with this course, along with an
+%% emphasis on expanding students' reasoning skills, will strengthen the
+%% connection between lower- and upper-division physics classes. Students
+%% will learn the importance of approximate calculations (Fermi
+%% estimates) that establish reasonable bounds on the exact
+%% answer. Students will learn that equations help us interpret the
+%% physical world, i.e. equations are used for much more than mindless
+%% algebraic manipulations.  Students will learn how integration and
+%% differentiation can be applied to real data - not just the perfect
+%% functions that are taught in math class. Students will learn how
+%% graphing of relationships is key tool for physicists. The utility of
+%% these skills will be firmly established by application to contemporary
+%% challenges.
 
@@
 
-\section{Outcomes}
-\emph{How has it gone?}
-
-\subsection{Faculty vote}
-\emph{Unanimous result.}
-\subsection{Transition year}
-\emph{Gradual transition, getting courses taught.}
+\section{Summary}
+\emph{Rearticulate big ideas. Shadowing}
 
 \section{Future research}
-\emph{Document learning trajectories, new courses.}
+\emph{Document learning trajectories, new courses.  pre-/post-,
+  understanding sense-making.}
 

### 2017-06-27 paper.tex
start tweaking intro to match 3 big ideas

@@
 \section{Introduction}
-Basic Paradigms in Physics ideas:
-\begin{enumerate}
-\item Attention to sequencing of concepts (and articulating content
-  learning goals)
-\item Faculty buy in and understanding of the sequencing and goals
-\item Active engagement
-\end{enumerate}
-\emph{This} is the Paradigms.  \emph{Explain why our 2.0 changes were
-  made, how they addressed these principles.  (i.e. sophomore-year
-  courses, math bits, giving students more electives as seniors) Also
-  articulate the process (both 1.0 and 2.0) in terms of these big
-  ideas.}
-
-The Paradigms in Physics program began around the time when our
-current students were born~\cite{manogue2001paradigms}.  The time has
-arrived for a change.  Over the years, we have identified a number of
-problems in our curriculum that we were unable to address without
-addressing the physics major curriculum as a whole.  In this paper, we
-share the motivation for our Paradigms 2.0 effort, the process by
-which we achieved consensus on an improved curriculum, and will give
-a summary and a few highlights of the new program.
+The Paradigms in Physics program is a rethinking of the upper-division
+physics curriculum that began around the time when our current
+students were born~\cite{manogue2001paradigms}.  The basic ideas
+behind the original Paradigms reform were (1) attention to the
+sequencing of concepts, (2) acheiving faculty buy-in, which requires
+faculty to understand the sequencing, and (3) implementation of active
+engagement through the upper-division curriculum.  In 2016, we began a
+rethinking of the Paradigms, with focus on the same principles guiding
+the original effort, but with the input of new faculty, and applying
+lessons learned over the past two decades.  In this paper, we will
+explain the reasoning behind this Paradigms 2.0 effort, the process by
+which we achieved developed and reached consensus on an improved
+curriculum, and will give both a summary and a few highlights of the
+new program.
 
@@
 There were a number of factors that motivated us to embark on the
-Paradigms~2.0 process.  The original Paradigms in Physics curricular
-reform from 200? was showing its age.  In the intervening years we
-have made a large number of modifications, such as reordering
-Paradigms, introducing one new Paradigm, and introducing and changing
-computational physics courses.
+Paradigms~2.0 process.  In the last two decades, we have observed
+challenges students face in the Paradigms.  While in many cases we
+addressed these by changing and reordering the courses, we felt a look
+at the entire curriculum was in order.  In addition, our faculty
+buy-in and aware was diminishing by attrition: while our younger
+faculty are enthusiastic about the Paradigms, many lack perspective on
+where students are in a given course, and what content is essential
+for students to grasp prior to a following course.
 
@@
 the fall of their junior year, and especially so for transfer
-students.  So we hoped to structure the courses such that non-transfer
+students.  So we aimed to structure the courses such that non-transfer
 students could be better prepared for their junior year, while

### 2017-06-27 paper.tex
fix typo

@@
 \section{Paradigms 2.0 process}
+We began the Paradigms 2.0 process in the Winter of 2016, having in
+the previous year obtained faculty agreement, and support from our
+Department head.  The process was spearheaded by a committee of four
+(the authors DR, LG, EM, and CAM, with EvZ present at meetings
+documenting our process).  During the Winter quarter, we informed the
+community of our process through a colloquium, and collected student
+and faculty perspectives on the existing curriculum.  We then
+interviewed faculty who recently taught each of our existing courses
+to document topics that were currently covered.  This resulted in a
+total of XXX index cards in YYY stacks, corresponding to courses, with
+cards color coded by broad topic.
+
+The committee met for ??? weeks to discuss existing challenges and
+sort the cards into new stacks corresponding to new and reordered
+courses.  When we discussed major changes to a course, we often
+invited interested faculty to join us to provide their own perspective
+on possible challenges and improvements.  Once we had a draft
+proposal, we began inviting each faculty member in to see the cards
+and discuss the proposed sequence of courses.  In addition, we had a
+meeting with all the current students to explain the proposed changes
+and request feedback.
+
+When we had obtained feedback from the entire faculty, we scheduled
+\emph{two} faculty meetings (DURATION?).  In the first meeting, we
+presented our proposal in detail.  This was needed for several
+reasons.  Those faculty we met with first may not have seen the final
+proposal.  Some faculty during their individual meetings chose to
+focus on a small subset of the curriculum, often relating to courses
+they had themselves taught.  In one meeting we presented the proposal
+and fielded questions.  In a second faculty meeting, we again
+presented our proposal, and
+
 \emph{How did we do it?}
@@
 We propose a course that explicitly focuses on physics majors'
-mathematical fluency and sense- making skills. Introductory courses
+mathematical fluency and sense-making skills. Introductory courses
 begin to develop these aspects, but the primary goals of those courses
@@
 
-This course will be taken just after the introductory physics
-sequence. It will leverage students' understandings of simple
-mechanical systems from introductory physics to develop their
-mathematical fluency in the context of less idealized, more
-sophisticated systems. Each physics topic (velocity-dependent forces,
-Lagrangian mechanics, Hamiltonian mechanics, special relativity) will
-focus on specific math methods (e.g. differential equations, calculus
-of variations, generalized coordinates, Legendre transformations,
-linear transformations, hyperbola geometry) and will explicitly
-feature strategies to make physical sense of mathematics (symbolic
-problem solving, reading equations like sentences, making
-idealization, coordinating representations, limiting case analysis,
-dimensional checking).
+This course is taken either just after or concurrently with the last
+quarter of the introductory physics sequence. It leverages students'
+understandings of simple mechanical systems from introductory physics
+to develop their mathematical fluency in the context of less
+idealized, more sophisticated systems. Each physics topic
+(velocity-dependent forces, Lagrangian mechanics, Hamiltonian
+mechanics, special relativity) will focus on specific math methods
+(e.g. differential equations, calculus of variations, generalized
+coordinates, Legendre transformations, linear transformations,
+hyperbola geometry) and will explicitly feature strategies to make
+physical sense of mathematics (symbolic problem solving, reading
+equations like sentences, making idealization, coordinating
+representations, limiting case analysis, dimensional checking).
 
@@
 
-\section{Future research}
-\emph{Document learning trajectories, new courses.  pre-/post-,
+\emph{Future research: Document learning trajectories, new courses.  pre-/post-,
   understanding sense-making.}
@@
 
-\section{Conclusions}
-\emph{We made some changes, and the process went well.}
-
-\acknowledgments{Put references below the acknowledgements (and appendixes, if any).}
+\acknowledgments{This paper is based upon work supported by the
+  National Science Foundation under Grant No. 1323800 Supplement.}
 

### 2017-06-27 paper.tex
add paragraph from Liz

@@
 
-To physicists, mathematics is a language for telling stories about
-conceptual physical relationships.  Our understandings of conceptual
-connections guide us when we manipulate
-equations~\cite{gire2008resources, bing2007cognitive,
-  wittmann2015mathematical}[75].  For example, when solving ordinary
-differential equations by separating variables, we choose limits of
-integration so that the lower (and upper) limits for all variables
-describe the same physical state.  We monitor algebraic manipulations
-using conceptual understandings of physics quantities. For example, we
-make sure that each term in a sum has the same physical dimensions
-(e.g. you can't add a frequency to a frequency squared). We make sure
-that expressions on both sides of an equals sign have the same
-dimensions~\cite{lenz2016dimensional}. We consider limiting cases to
-make sure that a derived expression reasonably describes the system in
-extreme states~\cite{singh2002physical}. Our geometric understandings
-guide the formulation of equations~\cite{gire2014arrows}. Sometimes,
-we do not have adequate mathematics to describe complex physical
-systems. We use conceptual understandings to make choices about
-reasonable idealizations in order to create useful mathematical
-models~\cite{bing2007cognitive}.
+While the Challenges courses takes a more experimental/applied physics
+focus, the Methods of Theoretical Mechanics course has a more
+theoretical physics flavor. The theme of this course is the discussion
+of strategies for making sense of physics problems and symbolic
+problem solving. This sense-making includes coordinating and
+interpreting symbolic expressions with conceptual understandings,
+geometric relationships, and physical intuitions.
 
-Being fluent in this physics language involves both (1) translating a
-conceptual story into a mathematical representation and (2) discerning
-a conceptual story from a mathematical
-representation~\cite{tuminaro2007elements}. At the introductory level,
-students often see concepts and mathematics as being separate. These
-students do not expect equations to help their conceptual
-understanding~\cite{adams2006new, redish1998expectations} and have
-trouble formulating equations from a conceptual analysis. Students who
-do not have mathematical fluency for physics struggle in advanced
-physics courses~\cite{manogue2009cognitive}.
+This Theoretical Mechanics course is aimed at students who are taking
+or have completed the last course in the introductory sequence. The
+physics content of the course is advanced Newtonian mechanics,
+introduction to Lagrangian and Hamiltonian techniques, and special
+relativity. These topics at this stage for the students are convenient
+for explicit discussion of sense-making strategies in physics because
+(a) students are have recently studied problems that serve as limiting
+cases for more complex problems, (b) students at this level are
+transitioning from solving problems primarily involving numbers to
+problems with only symbolic parameters, and (c) we can discuss
+sense-making as strategies for approaching a problem, evaluating
+answers, and refining intuitions about relativistic and other
+unfamiliar situations. For example, we teach students how to use
+spacetime diagrams to develop a story about a relativistic situation
+and how to use hyperbola trig on these diagrams to perform Lorentz
+transformations that can be checked against the results of algebraic
+Lorentz transformations.  The sense-making goals of the course are on
+equal footing with the physics goals, and sense-making is integrated
+explicitly in all aspects of the course: exams, homework and during
+in-class activities. Our aim is that students will develop
+sense-making skills that will improve their learning in advanced
+courses and will impress their upper-division course instructors,
+research advisors, and future employers.
 
-We propose a course that explicitly focuses on physics majors'
-mathematical fluency and sense-making skills. Introductory courses
-begin to develop these aspects, but the primary goals of those courses
-tend to be for students to apply simple concepts (Newton's Laws,
-energy conservation, etc.) to relatively straightforward and very
-idealized problems.
+%% To physicists, mathematics is a language for telling stories about
+%% conceptual physical relationships.  Our understandings of conceptual
+%% connections guide us when we manipulate
+%% equations~\cite{gire2008resources, bing2007cognitive,
+%%   wittmann2015mathematical}[75].  For example, when solving ordinary
+%% differential equations by separating variables, we choose limits of
+%% integration so that the lower (and upper) limits for all variables
+%% describe the same physical state.  We monitor algebraic manipulations
+%% using conceptual understandings of physics quantities. For example, we
+%% make sure that each term in a sum has the same physical dimensions
+%% (e.g. you can't add a frequency to a frequency squared). We make sure
+%% that expressions on both sides of an equals sign have the same
+%% dimensions~\cite{lenz2016dimensional}. We consider limiting cases to
+%% make sure that a derived expression reasonably describes the system in
+%% extreme states~\cite{singh2002physical}. Our geometric understandings
+%% guide the formulation of equations~\cite{gire2014arrows}. Sometimes,
+%% we do not have adequate mathematics to describe complex physical
+%% systems. We use conceptual understandings to make choices about
+%% reasonable idealizations in order to create useful mathematical
+%% models~\cite{bing2007cognitive}.
 
-This course is taken either just after or concurrently with the last
-quarter of the introductory physics sequence. It leverages students'
-understandings of simple mechanical systems from introductory physics
-to develop their mathematical fluency in the context of less
-idealized, more sophisticated systems. Each physics topic
-(velocity-dependent forces, Lagrangian mechanics, Hamiltonian
-mechanics, special relativity) will focus on specific math methods
-(e.g. differential equations, calculus of variations, generalized
-coordinates, Legendre transformations, linear transformations,
-hyperbola geometry) and will explicitly feature strategies to make
-physical sense of mathematics (symbolic problem solving, reading
-equations like sentences, making idealization, coordinating
-representations, limiting case analysis, dimensional checking).
+%% Being fluent in this physics language involves both (1) translating a
+%% conceptual story into a mathematical representation and (2) discerning
+%% a conceptual story from a mathematical
+%% representation~\cite{tuminaro2007elements}. At the introductory level,
+%% students often see concepts and mathematics as being separate. These
+%% students do not expect equations to help their conceptual
+%% understanding~\cite{adams2006new, redish1998expectations} and have
+%% trouble formulating equations from a conceptual analysis. Students who
+%% do not have mathematical fluency for physics struggle in advanced
+%% physics courses~\cite{manogue2009cognitive}.
+
+%% We propose a course that explicitly focuses on physics majors'
+%% mathematical fluency and sense-making skills. Introductory courses
+%% begin to develop these aspects, but the primary goals of those courses
+%% tend to be for students to apply simple concepts (Newton's Laws,
+%% energy conservation, etc.) to relatively straightforward and very
+%% idealized problems.
+
+%% This course is taken either just after or concurrently with the last
+%% quarter of the introductory physics sequence. It leverages students'
+%% understandings of simple mechanical systems from introductory physics
+%% to develop their mathematical fluency in the context of less
+%% idealized, more sophisticated systems. Each physics topic
+%% (velocity-dependent forces, Lagrangian mechanics, Hamiltonian
+%% mechanics, special relativity) will focus on specific math methods
+%% (e.g. differential equations, calculus of variations, generalized
+%% coordinates, Legendre transformations, linear transformations,
+%% hyperbola geometry) and will explicitly feature strategies to make
+%% physical sense of mathematics (symbolic problem solving, reading
+%% equations like sentences, making idealization, coordinating
+%% representations, limiting case analysis, dimensional checking).
 

### 2017-06-27 paper.tex
work on summary

@@
 and fielded questions.  In a second faculty meeting, we again
-presented our proposal, and
-
-\emph{How did we do it?}
-
-\emph{Faculty vote}
+presented our proposal, and opened the floor for discussion.  After
+considerable discussion, we unanimously voted to adopt the proposed
+changes.
 
@@
 \section{Summary}
-\emph{Rearticulate big ideas. Shadowing}
+We have developed a significant incremental change to the Paradigms
+curriculum.  This change focused on the first two big ideas underlying
+the Paradigms: (1) close attention to content ordering, and (2)
+developing a consensus and understanding of the curriculum among our
+faculty members.  Our commitment to using active engagement in our
+upper-division curriculum remains unchanged.  While the specific
+content ordering we developed is interesting, the \emph{process} by
+which we reached it, and developed a faculty consensus is at the
+essence of what makes distinctive the Paradigms in Physics program.
+The process of developing faculty understanding of the Paradigms is
+ongoing, as we are having professors new to the Paradigms shadow
+experienced faculty as they teach in the Paradigms, prior to teaching
+the courses themselves..
 
-\emph{Future research: Document learning trajectories, new courses.  pre-/post-,
-  understanding sense-making.}
+Future work will involve documenting the learning trajectories that we
+have developed in our curriculum.  Furthermore, we are engaging in a
+project to study our students' development of sense-making skills in
+particular through the two new courses that are intended specifically
+to ramp up those skills (instincts?).
+
+We look forward to another two decades of the Paradigms!
 
@@
 
-\acknowledgments{This paper is based upon work supported by the
-  National Science Foundation under Grant No. 1323800 Supplement.}
+\acknowledgments{This work was supported by the National Science
+  Foundation under Grant No. 1323800 Supplement.}
 

### 2017-06-27 paper.tex
count cards

@@
 to document topics that were currently covered.  This resulted in a
-total of XXX index cards in YYY stacks, corresponding to courses, with
-cards color coded by broad topic.
+total of $\sim$700 index cards in $\sim$30 stacks, with each stack
+corresponding to one course, and each card describing a topic,
+color-coded as in Fig.~\ref{fig:schedule}.
 

### 2017-06-27 paper.tex
polish process text a tad

@@
 must be able to graduate with two years of upper-division
-courses~\cite{manogue2001paradigms}.  The process of redesigning the
-curriculum was dominated by the creation index cards listing subject
-content, and sorting these cards into courses.  This process involved
-the entire faculty, and culminated in a unanimous agreement to adopt
-the resulting curriculum.
+courses~\cite{manogue2001paradigms}.  On top of this, there was a
+desire to provide students with a broad knowledge of physics prior to
+the GRE exams in the fall of their senior year.  The process of
+redesigning the curriculum was dominated by the creation index cards
+listing subject content, and sorting these cards into courses.  This
+process involved the entire faculty, and culminated in a unanimous
+agreement to adopt the resulting curriculum.
 
@@
 
-The committee met for ??? weeks to discuss existing challenges and
-sort the cards into new stacks corresponding to new and reordered
-courses.  When we discussed major changes to a course, we often
-invited interested faculty to join us to provide their own perspective
-on possible challenges and improvements.  Once we had a draft
-proposal, we began inviting each faculty member in to see the cards
-and discuss the proposed sequence of courses.  In addition, we had a
-meeting with all the current students to explain the proposed changes
-and request feedback.
+The committee met twice a week to discuss existing challenges and sort
+the cards into new stacks corresponding to new and reordered courses.
+When we discussed major changes to a course, we often invited
+interested faculty to join us to provide their own perspective on
+possible challenges and improvements.  Once we had a draft proposal,
+we began inviting each faculty member in to see the cards and discuss
+the proposed sequence of courses.  In addition, we had a meeting with
+all the current students to explain the proposed changes and request
+feedback.
 
 When we had obtained feedback from the entire faculty, we scheduled
-\emph{two} faculty meetings (DURATION?).  In the first meeting, we
-presented our proposal in detail.  This was needed for several
-reasons.  Those faculty we met with first may not have seen the final
-proposal.  Some faculty during their individual meetings chose to
-focus on a small subset of the curriculum, often relating to courses
-they had themselves taught.  In one meeting we presented the proposal
-and fielded questions.  In a second faculty meeting, we again
-presented our proposal, and opened the floor for discussion.  After
-considerable discussion, we unanimously voted to adopt the proposed
-changes.
+\emph{two} full faculty meetings.  In the first meeting, we presented
+our proposal in detail and invited questions, but not discussion.
+This was needed for several reasons.  Those faculty we met with first
+may not have seen the final proposal.  Some faculty during their
+individual meetings chose to focus on a small subset of the
+curriculum, often relating to courses they had themselves taught.  In
+one meeting we presented the proposal and fielded questions.  In a
+second faculty meeting, we again presented our proposal, and opened
+the floor for discussion.  After considerable discussion, we
+unanimously voted to adopt the proposed changes.
 

### 2017-06-27 paper.tex
count facutly members

@@
 \paragraph{New faculty}
-At this stage, (XX) of our faculty was not present during the orignal
-Paradigms effort.  Many of the newer faculty have only a superficial
-undertanding of our course structure.  This poses a challenge when
-these professors teach courses that are closely intertwined with one
-another.
+% Corinne, Henri, David McI, Janet, Tom = 5
+
+% David R, Ethan, Oksana, Bo, Guenter, Weihong, Matt, Yun-Shik, Heidi,
+% Liz, Davide = 11
+Only 5 of our 16 current tenure-line faculty were involved in the
+orignal Paradigms effort.  Some of the newest faculty have only a
+superficial undertanding of our course structure.  This poses a
+challenge when these professors teach courses that are closely
+intertwined with one another.
 

### 2017-06-28 paper.tex
Update on Overleaf.

@@
 \pagestyle{empty}		% page numbers added later, when compiling the whole proceedings
+
+\newcommand\corinnesays[1]{\textcolor{red}{[\sc CAM: \em{#1}]}}
+
 \begin{document}
@@
 \affiliation{Department of Physics, Oregon State University, Corvallis, Oregon, 97331}
-\author{Tevian Dray}
-\affiliation{Department of Mathematics, Oregon State University, Corvallis, Oregon, 97331}
 \author{Corinne A. Manogue}
@@
 process to revise our Paradigms in Physics curriculum for physics
-majors.  This poster will describe both the process by which our
+majors.  This paper will describe both the process by which our
 department achieved consensus on this curricular change, and the
@@
 \maketitle
-
+% * <ethan.minot@oregonstate.edu> 2017-06-28T19:21:10.109Z:
+%
+% ^.
 \section{Introduction}
 The Paradigms in Physics program is a rethinking of the upper-division
-physics curriculum that began around the time when our current
-students were born~\cite{manogue2001paradigms}.  The basic ideas
-behind the original Paradigms reform were (1) attention to the
-sequencing of concepts, (2) acheiving faculty buy-in, which requires
-faculty to understand the sequencing, and (3) implementation of active
-engagement through the upper-division curriculum.  In 2016, we began a
-rethinking of the Paradigms, with focus on the same principles guiding
-the original effort, but with the input of new faculty, and applying
-lessons learned over the past two decades.  In this paper, we will
-explain the reasoning behind this Paradigms 2.0 effort, the process by
-which we achieved developed and reached consensus on an improved
-curriculum, and will give both a summary and a few highlights of the
-new program.
-
-\subsection{Background}
 The Paradigms in Physics project began in 1996, when three faculty
@@
 the GRE exams in the fall of their senior year.  The process of
-redesigning the curriculum was dominated by the creation index cards
+redesigning the curriculum was dominated by the creation of index cards
 listing subject content, and sorting these cards into courses.  This
@@
 
-A coupe of decades have passed since the original Paradigms effort.
+Two decades have passed since the original Paradigms effort.
 During this time we have made a number of changes: e.g. a new
@@
 the junior year, and a computational laboratory course was introduced
-to accompany the Paradigms.  We maintained the tradition of meeting
+to accompany the Paradigms.  The faculty maintained the tradition of meeting
 every three weeks to discuss issues relating to upper-division
 teaching.  New faculty arrived and learned to teach the new courses,
-and introduced their own approach to the courses.
+and introduced their own ideas to the courses.
 
@@
 Paradigms~2.0 process.  In the last two decades, we have observed
-challenges students face in the Paradigms.  While in many cases we
+a number of challenges students face in the Paradigms.  While in many cases we
 addressed these by changing and reordering the courses, we felt a look
 at the entire curriculum was in order.  In addition, our faculty
-buy-in and aware was diminishing by attrition: while our younger
+understanding of the details of the sequence was diminishing by attrition: 
+while our younger
 faculty are enthusiastic about the Paradigms, many lack perspective on
@@
 the fall of their junior year, and especially so for transfer
-students.  So we aimed to structure the courses such that non-transfer
+students.  So we aimed to structure and order the courses such that non-transfer
 students could be better prepared for their junior year, while
@@
 everything, without progress being made.  
+\corinnesays{This paragraph could perhaps be tightened.}
 
@@
 we began inviting each faculty member in to see the cards and discuss
-the proposed sequence of courses.  In addition, we had a meeting with
+the proposed sequence of courses.  In addition, we had a focus group with
 all the current students to explain the proposed changes and request
@@
 individual meetings chose to focus on a small subset of the
-curriculum, often relating to courses they had themselves taught.  In
-one meeting we presented the proposal and fielded questions.  In a
+curriculum, often relating to courses they had themselves taught.  In a
 second faculty meeting, we again presented our proposal, and opened
@@
 Finally, by reducing the number of Paradigms, we allow the curriculum
-to be taught with fewer faculty, albeit at a work load per professor.
+to be taught with fewer faculty, albeit at an increased work load per professor.
 
@@
 one formerly 4-week Paradigm (the one that consumed the extra week
-each quarter) or two 3-week Paradigms compressed.  The major
+each quarter), two 3-week Paradigms compressed, or three 3-week paradigms combined into two 5-week paradigms.  The major
 scheduling change is to place Static Fields (electrostatics and
-magnetostatics) in the spring quarter.  This intended to give students
+magnetostatics) in the spring quarter.  
+This intended to give students, especially transfer students, 
 a bit more time to take Vector Calculus.  It also puts all use of
 curvilinear coordinates in the spring quarter, which should ease the
-learning of Central Forces.  The final major change was the
-elimination of special relativity from the Paradigms, which is has now
-moved into the new sophomore-level Theoretical Mechanics course.
+learning of Central Forces.  The final major change was to move special relativity from the junior-year Paradigms to the new sophomore-level Theoretical Mechanics course.
 
@@
 posed a challenge regarding placement in the curriculum: it needed to
-be sufficiently late in order for students to be adequately prepared
+be sufficiently late for students to be adequately prepared
 for advanced mathematical content, but in order to be functional in
@@
 the students.
+\corinnesays{This paragraph is awkward.  I'm still thinking about some rewording.  This might be a good paragraph to discuss this afternoon.}
 
@@
 Electronics during the junior year (specifically as a prerequisite for
-Oscillations and Waves).  This is made possible by the reduction in
+the Oscillations and Waves Paradigm).  This is made possible by the reduction in
 Fall workload for incoming transfer students, who had previously
@@
 trouble-shooting skills prior to the in-class labs taught in
-Oscillations and Waves.
+Oscillations and Waves course.
 
@@
 levels and needs.  Transfer students now take computation alongside
-the non-transfer students, in a laboratory setting using pair
+the non-transfer students. The course is taught in a laboratory setting using pair
 programming to help new programmers to learn.
@@
 these courses ramp up student mathematical abilities prior to their
-junior year.  The Challenges course will focus on estimation,
+junior year.  The Challenges course focuses on estimation,
 dimensional reasoning, and interpretation of integrals, while
@@
 \subsubsection{Challenges in Contemporary Physics}
+\corinnesays{You might want to use your paragraph structure here, rather than sub-sub sections.}
 

### 2017-06-28 paper.tex
before changes with feedback from meeting

@@
 \newcommand\corinnesays[1]{\textcolor{red}{[\sc CAM: \em{#1}]}}
+\newcommand\davidsays[1]{\textcolor{blue}{[\sc DR: \em{#1}]}}
 
@@
 \author{David Roundy}
-\author{Liz Gire}
+\author{Elizabeth Gire}
 \author{Ethan Minot}
@@
 proposal, which was unanimously approved by the faculty.  Major
-changes include elimination or major changes to several courses (math
+changes include major changes to several courses (math
 methods, computational physics, modern physics, electronics, and
@@
 \paragraph{Transfer students}
-Our curriculum is strongly affected by our desire to accomodate
+Our curriculum is strongly affected by our desire to accommodate
 transfer students from community colleges.  This leads us to focus on
@@
 addressed without addressing the physics major curriculum as a whole,
-and any discussion of one tended to devolve into a discussion of
 everything, without progress being made.  
@@
 orignal Paradigms effort.  Some of the newest faculty have only a
-superficial undertanding of our course structure.  This poses a
+superficial understanding of our course structure.  This poses a
 challenge when these professors teach courses that are closely
@@
 Department head.  The process was spearheaded by a committee of four
-(the authors DR, LG, EM, and CAM, with EvZ present at meetings
+(the authors DR, EG, EM, and CAM, with EvZ present at meetings
 documenting our process).  During the Winter quarter, we informed the
@@
 students to receive the math instruction they require, and based on
-our previous experience with math-intensive weeks (the ``odd'' weeks)
+our previous experience with math-intensive weeks (the \davidsays{``extra''
+weeks having 3-week courses fitted into a 10-week quarter})
 in the Paradigms, we chose to integrate one week of ``math bits'' into
@@
 \subsection{Sophomore courses}
-We introduced two new sophomore-level courses: \emph{Challenges in
-  Contemporary Physics}, and \emph{Theoretical Mechanics}.  Both of
+We introduced two new sophomore-level courses: \emph{Physics of
+Contemporary Challenges} and \emph{Techniques of Theoretical Mechanics}.  Both of
 these courses ramp up student mathematical abilities prior to their
@@
 dimensional reasoning, and interpretation of integrals, while
-Theoretical Mechanics will teach power series approximations, and
-expose students to increased levels of mathematical sophistication.
+Theoretical Mechanics teaches power series approximations and
+exposes students to increased levels of mathematical sophistication.
 
-These courses \emph{will} have the challenge (which was removed from
-the computational course!) of teaching both juniors and sophomores
-together.  They will be taught in the winter and spring quarters, so
+These courses have the challenge (which was removed from
+the computational course!)\corinnesays{remove this parenthetical comment} 
+of teaching both juniors and sophomores
+together.  They are taught in the winter and spring quarters, so
 as to reduce the burden on transfer students in the Fall.  They were
@@
 discussed in class, with additional details offered in online video
-format. In summary, we emphasis the physics concepts and reasoning
+format. In summary, we emphasize the physics concepts and reasoning
 skills that allow professional physicists to quickly/quantitatively
@@
 help clarify the goal of each lecture. The weekly homework questions
-are directly related to contemporary challenges are require a variety
+are directly related to contemporary challenges requiring a variety
 of estimation techniques. The modern physics experiment, performed
@@
 
-While the Challenges courses takes a more experimental/applied physics
+While the Challenges course takes a more experimental/applied physics
 focus, the Methods of Theoretical Mechanics course has a more
@@
 
-This Theoretical Mechanics course is aimed at students who are taking
+The Theoretical Mechanics course is aimed at students who are taking
 or have completed the last course in the introductory sequence. The
@@
 (a) students are have recently studied problems that serve as limiting
-cases for more complex problems, (b) students at this level are
+cases for these more complex problems, (b) students at this level are
 transitioning from solving problems primarily involving numbers to
-problems with only symbolic parameters, and (c) we can discuss
-sense-making as strategies for approaching a problem, evaluating
+problems with only symbolic parameters, and (c) sense-making as strategies 
+can be discussed for approaching a problem, evaluating
 answers, and refining intuitions about relativistic and other
@@
 spacetime diagrams to develop a story about a relativistic situation
-and how to use hyperbola trig on these diagrams to perform Lorentz
+and how to use hyperbola trigonometry on these diagrams to perform Lorentz
 transformations that can be checked against the results of algebraic
 Lorentz transformations.  The sense-making goals of the course are on
-equal footing with the physics goals, and sense-making is integrated
+an equal footing with the physics goals, and sense-making is integrated
 explicitly in all aspects of the course: exams, homework and during
@@
 sense-making skills that will improve their learning in advanced
-courses and will impress their upper-division course instructors,
+courses and will be valued by their upper-division course instructors,
 research advisors, and future employers.
@@
 \section{Summary}
-We have developed a significant incremental change to the Paradigms
-curriculum.  This change focused on the first two big ideas underlying
+We have developed a significant change to the Paradigms
+curriculum.  This change focuses on the first two big ideas underlying
 the Paradigms: (1) close attention to content ordering, and (2)
@@
 upper-division curriculum remains unchanged.  While the specific
-content ordering we developed is interesting, the \emph{process} by
-which we reached it, and developed a faculty consensus is at the
-essence of what makes distinctive the Paradigms in Physics program.
+content ordering we developed may be interesting, the \emph{process} by
+which we reached it and developed a faculty consensus is at the
+essence of what makes the Paradigms in Physics program distinctive.
+Emily van Zee, a researcher with experience in ..., 
+\corinnesays{Emily, please add a short description of yourself} 
+has completed a retrospective study of the process we used to develop the 
+original Paradigms program and we have used her results extensively in 
+designing and implementing the Paradigms 2.0 process.  She has continued 
+to observe, study, and document our work as we have begun implementing 
+the revised curriculum. 
+\corinnesays{David:  Add a citation to http://physics.oregonstate.edu/portfolioswiki/toolkit:start}
 The process of developing faculty understanding of the Paradigms is
@@
 experienced faculty as they teach in the Paradigms, prior to teaching
-the courses themselves..
+the courses themselves.
 

### 2017-06-29 paper.tex

@@
 department achieved consensus on this curricular change, and the
-resulting curriculum.  The Paradigms 2.0 committee begain with a
+resulting curriculum.  The Paradigms 2.0 committee began with a
 survey of students and faculty, followed by individual interviews with
@@
 \section{Introduction}
-The Paradigms in Physics program is a rethinking of the upper-division
 The Paradigms in Physics project began in 1996, when three faculty
 members at Oregon State applied to the NSF for funding to redesign the
-upper-division curriculum.  The motivation for the original Paradigms
-project was in fact similar to the motivation for Paradigms 2.0: a
+upper-division curriculum.  The motivation for the original Paradigms~1.0
+project was in fact similar to the motivation for Paradigms~2.0: a
 desire to soften the ``brick wall'' encountered by students upon
@@
 
-The resulting curriculum was primarily composed of junior-year
-``Paradigms'' courses, followed by senior-year ``Capstone'' courses.
-The Paradigms were intensive 2-credit, 3-week courses which meet every
-day, for a total of seven hours per week.  Paradigms incorporate
+The resulting curriculum was primarily composed of intensive junior-year \emph{Paradigm}
+courses, followed by more conventional senior-year courses.
+The \emph{Paradigm} courses meet every day for a total of seven
+hours per week for 3 weeks.  These courses incorporate
 laboratory experiences and active engagement into the class, and
-typically have two problem sets per week.  Capstones were more
-traditional 3-credit courses which meet three hours per week for an
-entire quarter.  The Paradigms courses had integrated math content,
-followed by a Math Methods course.
+typically have two problem sets per week.  These courses also had
+integrated math content, and were followed by a \emph{Math Methods} course.
+The senior-year courses are more traditional 3-credit courses which meet
+three hours per week for an entire quarter.
 
-Two decades have passed since the original Paradigms effort.
+Two decades have passed since the original Paradigms~1.0 effort.
 During this time we have made a number of changes: e.g. a new
-\emph{first} Paradigm was introduced to soften the beginning of the
-junior year, the order of Paradigms was shuffled more than once, Math
-Methods was moved from the beginning of the senior year to the end of
+first \emph{Paradigm} course was introduced to soften the beginning of the
+junior year, the order of courses was shuffled more than once, \emph{Math
+Methods} was moved from the beginning of the senior year to the end of
 the junior year, and a computational laboratory course was introduced
-to accompany the Paradigms.  The faculty maintained the tradition of meeting
+to accompany the junior-year \emph{Paradigm} courses.  The faculty maintained the tradition of meeting
 every three weeks to discuss issues relating to upper-division
@@
 Paradigms~2.0 process.  In the last two decades, we have observed
-a number of challenges students face in the Paradigms.  While in many cases we
+a number of challenges students face in our major.  While in many cases we
 addressed these by changing and reordering the courses, we felt a look
@@
 while our younger
-faculty are enthusiastic about the Paradigms, many lack perspective on
+faculty are enthusiastic about the Paradigms program, many lack perspective on
 where students are in a given course, and what content is essential
@@
 Only 5 of our 16 current tenure-line faculty were involved in the
-orignal Paradigms effort.  Some of the newest faculty have only a
+orignal Paradigms~1.0 effort.  Some of the newest faculty have only a
 superficial understanding of our course structure.  This poses a
@@
 \section{Paradigms 2.0 process}
-We began the Paradigms 2.0 process in the Winter of 2016, having in
+We began the Paradigms~2.0 process in the Winter of 2016, having in
 the previous year obtained faculty agreement, and support from our
@@
 new sophomore-level courses (which may be taken in the junior year) to
-better prepare our majors for the Paradigms.  We eliminated Modern
-Physics and the Classical Mechanics capstone in favor of the two new
-sophomore-level courses.  We eliminated the Math Methods course in
-favor of ``math bits'' integrated into the Paradigms themselves.
-Finally, we restructured our nine 3-week Paradigms courses into six
-5-week Paradigms.  We changed our computational physics requirement,
+better prepare our majors for the their junior year.  We eliminated \emph{Modern
+Physics} and the \emph{Classical Mechanics Capstone} in favor of the two new
+sophomore-level courses.  We eliminated the \emph{Math Methods} course in
+favor of ``\emph{Math Bits}'' integrated into the \emph{Paradigm} courses.
+Finally, we restructured our nine 3-week \emph{Paradigm} courses into six
+5-week \emph{Paradigm} courses.  We changed our computational physics requirement,
 and reduced the number of electronics courses.
 
-\subsection{5-week Paradigms}
+\subsection{5-week \emph{Paradigm} courses}
 To begin with the most distinctive feature of our curriculum, we
@@
 the junior year, which we have found effective.  However, we chose to
-change from three 3-week Paradigms per quarter to two 5-week Paradigms
+change from three 3-week courses per quarter to two 5-week courses
 per quarter.    This change has several advantages.  It gives students
@@
 gives a bit more flexibility, for instance, in case of illness.
-Finally, by reducing the number of Paradigms, we allow the curriculum
+Finally, by reducing the number of courses, we allow the curriculum
 to be taught with fewer faculty, albeit at an increased work load per professor.
@@
 We put considerable thought into the content and ordering of the new
-Paradigms.  In most cases, we think of each 5-week Paradigm as either
-one formerly 4-week Paradigm (the one that consumed the extra week
-each quarter), two 3-week Paradigms compressed, or three 3-week paradigms combined into two 5-week paradigms.  The major
-scheduling change is to place Static Fields (electrostatics and
+courses.  In most cases, we think of each 5-week \emph{Paradigm} course as either
+one formerly 4-week \emph{Paradigm} course (the one that consumed the extra week
+each quarter), two 3-week courses compressed, or three 3-week courses combined
+into two 5-week paradigms.  The major
+scheduling change is to place \emph{Static Fields} (electrostatics and
 magnetostatics) in the spring quarter.  
-This intended to give students, especially transfer students, 
-a bit more time to take Vector Calculus.  It also puts all use of
+This gives students, especially transfer students, 
+a bit more time to take \emph{Vector Calculus}.  It also puts all use of
 curvilinear coordinates in the spring quarter, which should ease the
-learning of Central Forces.  The final major change was to move special relativity from the junior-year Paradigms to the new sophomore-level Theoretical Mechanics course.
+learning of \emph{Central Forces}.  The final major change was to move
+special relativity from the junior year to the new sophomore-level
+\emph{Theoretical Mechanics} course.
 
 \subsection{Math bits}
-We chose to eliminate the Math Methods course in favor of formal
-``math bits'' in the Paradigms.  Our previous Math Methods Capstone
+We chose to eliminate the \emph{Math Methods} course in favor of formal
+\emph{Math Bits} in the \emph{Paradigm} courses.  Our previous \emph{Math Methods} course
 posed a challenge regarding placement in the curriculum: it needed to
@@
 advanced math methods, we chose to integrate the needed math content
-into the Paradigms, and allow advanced students to take our
-graduate-level Math Methods course.  In order to ensure that our
+into the \emph{Paradigm} courses, and allow advanced students to take our
+graduate-level \emph{Math Methods} course.  In order to ensure that our
 students to receive the math instruction they require, and based on
@@
 weeks having 3-week courses fitted into a 10-week quarter})
-in the Paradigms, we chose to integrate one week of ``math bits'' into
-each Paradigm.  The math bits for the entire year are treated as a
+in the \emph{Paradigm} courses, we chose to integrate one week of \emph{Math Bits} into
+each \emph{Paradigm} course.  The \emph{Math Bits} for the entire year are treated as a
 distinct teaching assignment, so students will be accustomed to the
 treatment of math as a distinct subject supporting the physics
-content.  This provides some relief from teaching for the Paradigms
-professors, and at the same time ensures that they do not succumb to
-the temptation to short-change the math content to the detriment of
-the students.
+content.  This provides some relief from teaching for
+professors teaching \emph{Paradigm} courses, and at the same time ensures
+that they do not succumb to the temptation to short-change the math content
+to the detriment of the students.
 \corinnesays{This paragraph is awkward.  I'm still thinking about some rewording.  This might be a good paragraph to discuss this afternoon.}
@@
 \subsection{Electronics}
-We chose to reduce the Electronics requirement from two 3-credit
+We chose to reduce the \emph{Electronics} requirement from two 3-credit
 courses to one 3-credit course.  For many physics majors, this is more
@@
 A final change to Electronics is that we now will \emph{require}
-Electronics during the junior year (specifically as a prerequisite for
-the Oscillations and Waves Paradigm).  This is made possible by the reduction in
+\emph{Electronics} during the junior year (specifically as a prerequisite for
+\emph{Oscillations and Waves}).  This is made possible by the reduction in
 Fall workload for incoming transfer students, who had previously
-usually taken Electronics as a senior.  This change has enabled us to
+usually taken \emph{Electronics} as a senior.  This change has enabled us to
 articulate distinct and sequenced learning outcomes from these two
@@
 trouble-shooting skills prior to the in-class labs taught in
-Oscillations and Waves course.
+\emph{Oscillations and Waves}.
 
@@
 Over the last six years, we have been developing a 1-credit
-computational laboratory course that accompanies the Paradigms.  We
+computational laboratory course that accompanies the \emph{Paradigm} courses.  We
 chose to require this course, while removing a requirement for a
@@
 these courses ramp up student mathematical abilities prior to their
-junior year.  The Challenges course focuses on estimation,
+junior year.  The \emph{Challenges} course focuses on estimation,
 dimensional reasoning, and interpretation of integrals, while
-Theoretical Mechanics teaches power series approximations and
+\emph{Theoretical Mechanics} teaches power series approximations and
 exposes students to increased levels of mathematical sophistication.
 
-These courses have the challenge (which was removed from
-the computational course!)\corinnesays{remove this parenthetical comment} 
-of teaching both juniors and sophomores
+These courses have the challenge of teaching both juniors and sophomores
 together.  They are taught in the winter and spring quarters, so
@@
 explicitly constructed to \emph{not} teach any content required for
-Fall or Winter Paradigms so that they can be taken concurrently with
-the Paradigms.
+Fall or Winter junior-year courses so that they can be taken concurrently with
+those courses.
 
-\subsubsection{Challenges in Contemporary Physics}
+\subsubsection{Physics of Contemporary Challenges}
 \corinnesays{You might want to use your paragraph structure here, rather than sub-sub sections.}
 
-In ``The Physics of Contemporary Challenges'' we prepare students to
+In \emph{The Physics of Contemporary Challenges} we prepare students to
 apply physics concepts and physical reasoning skills to sustainable
@@
 
-While the Challenges course takes a more experimental/applied physics
-focus, the Methods of Theoretical Mechanics course has a more
+While the \emph{Challenges} course takes a more experimental/applied physics
+focus, the \emph{Techniques of Theoretical Mechanics} course has a more
 theoretical physics flavor. The theme of this course is the discussion
@@
 
-The Theoretical Mechanics course is aimed at students who are taking
+The \emph{Theoretical Mechanics} course is aimed at students who are taking
 or have completed the last course in the introductory sequence. The
@@
 curriculum.  This change focuses on the first two big ideas underlying
-the Paradigms: (1) close attention to content ordering, and (2)
+the Paradigms project: (1) close attention to content ordering, and (2)
 developing a consensus and understanding of the curriculum among our
@@
 original Paradigms program and we have used her results extensively in 
-designing and implementing the Paradigms 2.0 process.  She has continued 
+designing and implementing the Paradigms~2.0 process.  She has continued 
 to observe, study, and document our work as we have begun implementing 
@@
 The process of developing faculty understanding of the Paradigms is
-ongoing, as we are having professors new to the Paradigms shadow
-experienced faculty as they teach in the Paradigms, prior to teaching
-the courses themselves.
+ongoing, as we are having professors new to the department shadow
+experienced faculty as they teach \emph{Paradigm} course, prior to teaching
+the same course themselves.
 

### 2017-06-29 paper.tex

@@
 
-\newcommand\corinnesays[1]{\textcolor{red}{[\sc CAM: \em{#1}]}}
-\newcommand\davidsays[1]{\textcolor{blue}{[\sc DR: \em{#1}]}}
+\newcommand\corinnesays[1]{\textcolor{red}{[\sc CAM: {\em#1}]}}
+\newcommand\davidsays[1]{\textcolor{blue}{[\sc DR: {\em#1}]}}
 
@@
 \section{Introduction}
+\davidsays{Why does a reader care? A dramatic example of curriculum change.
+A dramatic revision of said.  So others can adopt this method of change.  This is
+an implementation of education research.  It is an exemplar for the community.
+}
 The Paradigms in Physics project began in 1996, when three faculty
@@
 addressed these by changing and reordering the courses, we felt a look
-at the entire curriculum was in order.  In addition, our faculty
+at the entire curriculum was in order.  \davidsays{Triangulation of different sources.  Where did we get info about this issue?}
+In addition, our faculty
 understanding of the details of the sequence was diminishing by attrition: 
@@
 
-When we had obtained feedback from the entire faculty, we scheduled
-\emph{two} full faculty meetings.  In the first meeting, we presented
-our proposal in detail and invited questions, but not discussion.
-This was needed for several reasons.  Those faculty we met with first
-may not have seen the final proposal.  Some faculty during their
+After incorporating the feedback from  individual meetings with every
+faculty member, we scheduled
+two full faculty meetings with a week in between.  In the first meeting,
+we presented our proposal in detail and invited questions, but not discussion.
+This was needed for a couple of reasons:  those faculty we met with first
+may not have seen the final proposal, and some faculty during their
 individual meetings chose to focus on a small subset of the
-curriculum, often relating to courses they had themselves taught.  In a
-second faculty meeting, we again presented our proposal, and opened
-the floor for discussion.  After considerable discussion, we
-unanimously voted to adopt the proposed changes.
+curriculum, often relating to courses they had themselves taught.
+During the following week faculty engaged in hallway discussions of the proposal.
+This week helped ensure that the faculty did not feel rushed into a vote.
+In the second faculty meeting, we again presented our proposal, and opened
+the floor for discussion.  After considerable discussion, largely on
+changes that were not part of the proposal, we unanimously voted to adopt
+the proposed changes.
 
@@
   with shadows.\label{fig:schedule}}
+  \davidsays{Make years more clear, and avoid duplicating numbers.  Braces might be nice.
+  Fix arrow from Theoretical Mechanics to Central Forces.  Cut out the shadows.
+  Add borders (other depts, required, and elective).  Consider showing the transfer
+  student distinction within each year, possibly with a little gap.  Maybe switch to single
+  column, either rotated or with Fall/Winter/Spring horizontally oriented.
+  Check if bigger (proportional) math courses will work.}
 \end{figure*}
@@
 \subsection{5-week \emph{Paradigm} courses}
+\davidsays{Advising flexibility, student focus, cohort building---daily,
+active engagement.}
 To begin with the most distinctive feature of our curriculum, we
@@
 change from three 3-week courses per quarter to two 5-week courses
-per quarter.    This change has several advantages.  It gives students
+per quarter.    This change has several advantages.  \davidsays{Allows opportunity for
+formative assessment and feedback for students.  Gives a chance for students to
+catch up who fall behind.  Time to adjust to a new professor.  Simplifies teaching
+loads} It gives students
 a bit more time with each professor prior to their final exam.  This
@@
 \subsection{Math bits}
+\davidsays{Discuss extra week and previous experience earlier.}
 We chose to eliminate the \emph{Math Methods} course in favor of formal

### 2017-06-29 paper.tex
comments from Mike

@@
 classical mechanics), including the introduction of two sophomore-year
-courses designed specifically to help prepare students to for their
+courses designed specifically to help prepare students for their
 upper-division courses.
@@
 graduate-level \emph{Math Methods} course.  In order to ensure that our
-students to receive the math instruction they require, and based on
+students receive the math instruction they require, and based on
 our previous experience with math-intensive weeks (the \davidsays{``extra''
@@
 the non-transfer students. The course is taught in a laboratory setting using pair
-programming to help new programmers to learn.
+programming to help new programmers to learn. \davidsays{CITE!}
 
@@
 \emph{Theoretical Mechanics} teaches power series approximations and
-exposes students to increased levels of mathematical sophistication.
+exposes students to increased levels of mathematical sophistication. \davidsays{ADD CITATIONS SOMEWHERE HERE TO KELBY AND MACKENZIE}
 
@@
 experimental physics---in each case motivated by one or more
-``contemporary challenges''. Together with new physics concepts,
+``contemporary challenges.'' Together with new physics concepts,
 students are given new reasoning tools, such as the equipartition
@@
 explicitly in all aspects of the course: exams, homework and during
-in-class activities. Our aim is that students will develop
+in-class activities. \davidsays{CITE HERE?} Our aim is that students will develop
 sense-making skills that will improve their learning in advanced

### 2017-06-29 paper.tex
more edits to vertical figure
everyone's comments accounted for

@@
 The senior-year courses are more traditional 3-credit courses which meet
-three hours per week for an entire quarter.
+three hours per week for an entire 10-week quarter.
 
@@
 where students are in a given course, and what content is essential
-for students to grasp prior to a following course.
+for students to grasp for a subsequent course.
 
@@
 Our curriculum is strongly affected by our desire to accommodate
-transfer students from community colleges.  This leads us to focus on
+transfer students from community colleges, who make up about a \davidsays{quarter}
+of our Physics majors.  This leads us to focus on
 a physics major in which all upper-division courses are taken in two
@@
 students could be better prepared for their junior year, while
-transfer students could postpone a few courses so that their junior
-Fall would be no more difficult than that of any other student.
+transfer students could postpone a few courses so that the
+Fall quarter of their junior year would be no more difficult than that of any other student.
 
 \paragraph{Coupled curricular changes}
-A backlog of curricular changes have accumulated that could not be
-separated from an examination of the curriculum as a whole.  Most
-notable were discussions of the requirements for computational
-physics, the amount of required electronics credits, and the role of
-our modern physics course.  These persistent issues were not easily
-addressed without addressing the physics major curriculum as a whole,
-everything, without progress being made.  
-\corinnesays{This paragraph could perhaps be tightened.}
+A backlog of curricular changes had accumulated that could not be
+separated from an examination of the curriculum as a whole.  The requirements
+for computational physics, the number of required electronics credits, and the
+role of our modern physics course could not easily be addressed separately.
 
@@
 Only 5 of our 16 current tenure-line faculty were involved in the
-orignal Paradigms~1.0 effort.  Some of the newest faculty have only a
+original Paradigms~1.0 effort.  Some of the newest faculty have only a
 superficial understanding of our course structure.  This poses a
@@
 a bit more time with each professor prior to their final exam.  This
-gives a bit more flexibility, for instance, in case of illness.
+gives more flexibility, for instance, in case of illness.
 Finally, by reducing the number of courses, we allow the curriculum
@@
 \emph{Oscillations and Waves}).  This is made possible by the reduction in
-Fall workload for incoming transfer students, who had previously
+Fall workload for incoming transfer students, who had
 usually taken \emph{Electronics} as a senior.  This change has enabled us to
@@
 \emph{Theoretical Mechanics} teaches power series approximations and
-exposes students to increased levels of mathematical sophistication. \davidsays{ADD CITATIONS SOMEWHERE HERE TO KELBY AND MACKENZIE}
+exposes students to increased levels of mathematical sophistication and
+sense-making strategies. \davidsays{ADD CITATIONS SOMEWHERE HERE TO KELBY AND MACKENZIE}
 
@@
 retain as many potential physics majors as possible.
+\davidsays{Add in reference and note?}
 
@@
 introduction to Lagrangian and Hamiltonian techniques, and special
-relativity. These topics at this stage for the students are convenient
+relativity. These topics are convenient
 for explicit discussion of sense-making strategies in physics because
-(a) students are have recently studied problems that serve as limiting
+(a) these students have recently studied problems that serve as limiting
 cases for these more complex problems, (b) students at this level are
@@
 particular through the two new courses that are intended specifically
-to ramp up those skills (instincts?).
+to ramp up those skills.
 

### 2017-06-29 paper.tex
fix gray area
cut horizontal figure

@@
 
-\begin{figure*}
-\includegraphics[width=\textwidth]{schedule}
-\caption{Hypothetical student schedules for (a) a student entering
-  without calculus and (b) a transfer student.  The Dependencies
-  between courses are noted with arrows.  The area of each course is
-  proportional to the credit value of the course, with the exception
-  of math courses, which are shown in small boxes even though they are
-  4-credit courses.  The two newly introduced courses are indicated
-  with shadows.\label{fig:schedule}}
-  \davidsays{Make years more clear, and avoid duplicating numbers.  Braces might be nice.
-  Fix arrow from Theoretical Mechanics to Central Forces.  Cut out the shadows.
-  Add borders (other depts, required, and elective).  Consider showing the transfer
-  student distinction within each year, possibly with a little gap.  Maybe switch to single
-  column, either rotated or with Fall/Winter/Spring horizontally oriented.
-  Check if bigger (proportional) math courses will work.}
-\end{figure*}
-
+\begin{figure}
+\includegraphics[width=\columnwidth]{vertical-schedule}
+\caption{Hypothetical student schedule in the new curriculum.
+  Transfer students begin in year 3, and take the marked additional
+  courses.\label{fig:schedule}}
+\end{figure}
 \section{Changes made}
@@
 \subsection{Math bits}
-\davidsays{Discuss extra week and previous experience earlier.}
-We chose to eliminate the \emph{Math Methods} course in favor of formal
-\emph{Math Bits} in the \emph{Paradigm} courses.  Our previous \emph{Math Methods} course
-posed a challenge regarding placement in the curriculum: it needed to
-be sufficiently late for students to be adequately prepared
-for advanced mathematical content, but in order to be functional in
-the curriculum it should teach content that is required by other
-physics courses, and thus should be earlier in the curriculum.  In
-response to these issues, and in recognition that the majority of our
-majors are not bound for graduate school, and have no need for
-advanced math methods, we chose to integrate the needed math content
-into the \emph{Paradigm} courses, and allow advanced students to take our
-graduate-level \emph{Math Methods} course.  In order to ensure that our
-students receive the math instruction they require, and based on
-our previous experience with math-intensive weeks (the \davidsays{``extra''
-weeks having 3-week courses fitted into a 10-week quarter})
-in the \emph{Paradigm} courses, we chose to integrate one week of \emph{Math Bits} into
-each \emph{Paradigm} course.  The \emph{Math Bits} for the entire year are treated as a
-distinct teaching assignment, so students will be accustomed to the
-treatment of math as a distinct subject supporting the physics
-content.  This provides some relief from teaching for
+Since Paradigms~1.0 was began, we introduced a
+math-intensive week preceding two of the \emph{Paradigm} courses.  These math weeks provided
+just-in-time preparation of the math skills required for those courses.
+We have found these math weeks to be effective and popular with students.  In contrast to
+these weeks the \emph{Math Methods} course is unwieldy, and challenging to place in
+the curriculum at a time where it is helpful to students.  Moreover, some of the content
+in \emph{Math Methods} was not actually required for our undergraduate courses, and only
+needed for students bound for graduate school.  We chose therefore to eliminate the \emph{Math Methods} course in favor of \emph{Math Bits} consisting of a single week of 
+just-in-time math content incorporated in each \emph{Paradigm} course.  Advanced
+students may also take our graduate-level \emph{Math Methods} course.  
+The \emph{Math Bits} for the entire year is taught by a single professor,
+which provides continuity and coherence for the students.  
+This provides some assistance for
 professors teaching \emph{Paradigm} courses, and at the same time ensures
@@
 to the detriment of the students.
-\corinnesays{This paragraph is awkward.  I'm still thinking about some rewording.  This might be a good paragraph to discuss this afternoon.}
 
@@
 
-\subsubsection{Physics of Contemporary Challenges}
-\corinnesays{You might want to use your paragraph structure here, rather than sub-sub sections.}
-
-In \emph{The Physics of Contemporary Challenges} we prepare students to
+\paragraph{Physics of Contemporary Challenges}
+In this course we prepare students to
 apply physics concepts and physical reasoning skills to sustainable
@@
 
-\subsubsection{Techniques of Theoretical Mechanics}
+\paragraph{Techniques of Theoretical Mechanics}
 
 While the \emph{Challenges} course takes a more experimental/applied physics
-focus, the \emph{Techniques of Theoretical Mechanics} course has a more
+focus, \emph{Theoretical Mechanics} has a more
 theoretical physics flavor. The theme of this course is the discussion
@@
 
-The \emph{Theoretical Mechanics} course is aimed at students who are taking
+This course is aimed at students who are taking
 or have completed the last course in the introductory sequence. The
@@
 to observe, study, and document our work as we have begun implementing 
-the revised curriculum. 
-\corinnesays{David:  Add a citation to http://physics.oregonstate.edu/portfolioswiki/toolkit:start}
+the revised curriculum~\cite{lessonslearned}.
 The process of developing faculty understanding of the Paradigms is
@@
 
-%% \begin{table*}[htbp]
-%% \caption{Hypothetical student schedule for a student entering without
-%%   calculus.  Students with AP Calculus can take the junior-year
-%%   courses one year earlier, and students who start calculus-based
-%%   physics in the Fall of their second year are not
-%%   delayed.\label{schedule}}
-%% \begin{ruledtabular}
-%% \begin{tabular}{rlll}
-%%   \textbf{Year} & \textbf{Fall} & \textbf{Winter} & \textbf{Spring} \\
-%%   \hline
-%%   1 & \mathcourse{Differential calculus}{251} &
-%%   \mathcourse{Integral calculus}{252} &
-%%   \mathcourse{Multivariable calculus}{254} \\
-%%     & & \onecredit{Freshman Seminar} & \fourcredit{Calc Physics I} \\
-%% \hline  2 & \mathcourse{Vector calculus}{255}
-%%   & \mathcourse{Differential equations}{256}
-%%   & \mathcourse{Linear algebra}{341} \\
-%%   & \fourcredit{Calc Physics II} & \fourcredit{Calc Physics III} &\mathcourse{Series \&
-%%     sequences}{253} \\
-%%   && \noted{Challenges}{3} & \noted{Theoretical Mechanics}{3}
-%%   \\
-%% \hline 3 & \paradigm{Quantum fundamentals} & \paradigm{Osciallations \& waves} & \paradigm{Static fields}
-%% \\
-%%   & \paradigm{Energy \& entropy} & \paradigm{Periodic systems} & \paradigm{Central forces}
-%% \\
-%%   & \onecredit{Computational lab I} & \onecredit{Computational lab II} & \onecredit{Computational lab III}
-%% \\
-%%   & \threecredit{Electronics} & \threecredit{Computer interfacing} & \onecredit{Research}
-%% \\
-%% \hline 4 & \capstone{Electromagnetism capstone} & \capstone{Quantum capstone} & \capstone{Thermal capstone}
-%% \\
-%% & \onecredit{Thesis} & \onecredit{Thesis} & \onecredit{Thesis}
-%% \\
-%% & \onecredit{Research} & \onecredit{Research}
-%% \end{tabular}
-%% \end{ruledtabular}
-%% \end{table*}
+\end{document}
 
-%% \begin{table*}[htbp]
-%% \caption{Hypothetical transfer student schedule.  Most transfer
-%%   students will have a heavier schedule in their transfer year due
-%%   to need to catch up on required math courses.\label{schedule}}
-%% \begin{ruledtabular}
-%% \begin{tabular}{rlll}
-%%   \textbf{Year} & \textbf{Fall} & \textbf{Winter} & \textbf{Spring} \\
-%%   \hline
-%% 3 & \paradigm{Quantum fundamentals} & \paradigm{Osciallations \& waves} & \paradigm{Static fields}
-%% \\
-%%   & \paradigm{Energy \& entropy} & \paradigm{Periodic systems} & \paradigm{Central forces}
-%% \\
-%%   & \onecredit{Computational lab I} & \onecredit{Computational lab II} & \onecredit{Computational lab III}
-%% \\
-%%   & \threecredit{Electronics} & \threecredit{Computer interfacing} & \onecredit{Research}
-%% \\
-%% & \mathcourse{Linear algebra}{341} & \noted{Challenges}{3} & \noted{Theoretical Mechanics}{3}
-%% \\
-%% &&\mathcourse{Vector calculus}{255}
-%% \\
-%% \hline 4 & \capstone{Electromagnetism capstone} & \capstone{Quantum capstone} & \capstone{Thermal capstone}
-%% \\
-%% & \onecredit{Thesis} & \onecredit{Thesis} & \onecredit{Thesis}
-%% \\
-%% & \onecredit{Research} & \onecredit{Research}
-%% \end{tabular}
-%% \end{ruledtabular}
-%% \end{table*}
-
-\end{document}
+\begin{figure*}
+\includegraphics[width=\textwidth]{schedule}
+\caption{Hypothetical student schedules for (a) a student entering
+  without calculus and (b) a transfer student.  The Dependencies
+  between courses are noted with arrows.  The area of each course is
+  proportional to the credit value of the course, with the exception
+  of math courses, which are shown in small boxes even though they are
+  4-credit courses.  The two newly introduced courses are indicated
+  with shadows.\label{fig:schedule}}
+  \davidsays{Make years more clear, and avoid duplicating numbers.  Braces might be nice.
+  Fix arrow from Theoretical Mechanics to Central Forces.  Cut out the shadows.
+  Add borders (other depts, required, and elective).  Consider showing the transfer
+  student distinction within each year, possibly with a little gap.  Maybe switch to single
+  column, either rotated or with Fall/Winter/Spring horizontally oriented.
+  Check if bigger (proportional) math courses will work.}
+\end{figure*}

### 2017-06-29 paper.tex
further refinement

@@
 process to revise our Paradigms in Physics curriculum for physics
-majors.  This paper will describe both the process by which our
-department achieved consensus on this curricular change, and the
-resulting curriculum.  The Paradigms 2.0 committee began with a
+majors.  We began with a
 survey of students and faculty, followed by individual interviews with
@@
 % ^.
-\section{Introduction}
-\davidsays{Why does a reader care? A dramatic example of curriculum change.
-A dramatic revision of said.  So others can adopt this method of change.  This is
-an implementation of education research.  It is an exemplar for the community.
-}
+%\section{Introduction}
+%\davidsays{Why does a reader care? A dramatic example of curriculum change.
+%A dramatic revision of said.  So others can adopt this method of change.  This is
+%an implementation of education research.  It is an exemplar for the community.
+%}
 The Paradigms in Physics project began in 1996, when three faculty
@@
 
-\subsection{Motivation for Paradigms 2.0}
-There were a number of factors that motivated us to embark on the
+%\subsection{Motivation for Paradigms 2.0}
+A number of factors motivated us to embark on the
 Paradigms~2.0 process.  In the last two decades, we have observed
@@
 addressed these by changing and reordering the courses, we felt a look
-at the entire curriculum was in order.  \davidsays{Triangulation of different sources.  Where did we get info about this issue?}
+at the entire curriculum was in order.
 In addition, our faculty
@@
 Our curriculum is strongly affected by our desire to accommodate
-transfer students from community colleges, who make up about a \davidsays{quarter}
-of our Physics majors.  This leads us to focus on
+transfer students from community colleges.  This leads us to focus on
 a physics major in which all upper-division courses are taken in two
@@
 sophomore-level courses.  We eliminated the \emph{Math Methods} course in
-favor of ``\emph{Math Bits}'' integrated into the \emph{Paradigm} courses.
+favor of \emph{Math Bits} integrated into the \emph{Paradigm} courses.
 Finally, we restructured our nine 3-week \emph{Paradigm} courses into six
@@
 \subsection{5-week \emph{Paradigm} courses}
-\davidsays{Advising flexibility, student focus, cohort building---daily,
-active engagement.}
 To begin with the most distinctive feature of our curriculum, we
 wanted to maintain the existing intensive 7-hour-per-week schedule in
-the junior year, which we have found effective.  However, we chose to
+the junior year, which we have found effective.  This schedule allows students to focus
+intensely on a single topic, it fosters the building of a cohort of students, and 
+the daily schedule is helpful for active engagement.  We chose to
 change from three 3-week courses per quarter to two 5-week courses
-per quarter.    This change has several advantages.  \davidsays{Allows opportunity for
-formative assessment and feedback for students.  Gives a chance for students to
-catch up who fall behind.  Time to adjust to a new professor.  Simplifies teaching
-loads} It gives students
-a bit more time with each professor prior to their final exam.  This
-gives more flexibility, for instance, in case of illness.
-Finally, by reducing the number of courses, we allow the curriculum
-to be taught with fewer faculty, albeit at an increased work load per professor.
+per quarter.  This gives students who fall behind a chance to catch up, and enables
+us to give more feedback to students prior to the final exam.  Having 3-credit courses
+simplifies the setting of teaching loads, and by reducing the number of courses, 
+we allow the curriculum to be taught with fewer faculty, albeit at an increased work
 

### 2017-06-29 paper.tex
a couple of citations added

@@
 the non-transfer students. The course is taught in a laboratory setting using pair
-programming to help new programmers to learn. \davidsays{CITE!}
+programming~\cite{mcdowell2006pair} to help new programmers to learn.
 
@@
 inclusion of the most engaging challenges, we aim to attract and
-retain as many potential physics majors as possible.
-\davidsays{Add in reference and note?}
+retain as many potential physics majors as 
+possible~\cite{kramer2016gathering}.
 

### 2017-06-29 paper.tex
Update on Overleaf.

@@
 students to dive deeper into a particular contemporary challenge. The
-term paper tests student's ability to make their own initial
+term paper tests students' ability to make their own initial
 assessment of a complex problem, and communicate their ideas with

### 2017-06-29 paper.tex

@@
 process to revise our Paradigms in Physics curriculum for physics
-majors.  We began with a
-survey of students and faculty, followed by individual interviews with
-the faculty teaching each existing course.  As we developed a plan to
+majors.  We began with a colloquium to inform the department of our
+plans and request their assistance, followed by a
+survey of students and faculty as well as individual interviews with
+the faculty teaching each course.  As we developed a plan to
 address student- and faculty-identified challenges in the curriculum,
@@
 the GRE exams in the fall of their senior year.  The process of
-redesigning the curriculum was dominated by the creation of index cards
+redesigning the curriculum was guided by the creation of index cards
 listing subject content, and sorting these cards into courses.  This
@@
 the floor for discussion.  After considerable discussion, largely on
-changes that were not part of the proposal, we unanimously voted to adopt
+changes that were not part of the proposal, the faculty unanimously voted to adopt
 the proposed changes.
-
-%\newcommand\mathcourse[2]{\emph{#1 (#2)}}
-\newcommand\mathcourse[2]{\emph{#1}}
-\newcommand\noted[2]{\textbf{#1} (#2)}
-\newcommand\paradigm[1]{{\sc #1} (3)}
-\newcommand\capstone[1]{#1 (3)}
-\newcommand\onecredit[1]{#1 (1)}
-\newcommand\threecredit[1]{#1 (3)}
-\newcommand\fourcredit[1]{#1 (4)}
 
@@
 simplifies the setting of teaching loads, and by reducing the number of courses, 
-we allow the curriculum to be taught with fewer faculty, albeit at an increased work
+we allow the curriculum to be taught with fewer faculty, albeit at an increased
+work load per professor.
 

### 2017-06-29 paper.tex
first submitted draft

@@
 \newcommand\davidsays[1]{\textcolor{blue}{[\sc DR: {\em#1}]}}
+\renewcommand\davidsays[1]{}
 
@@
 \section{Paradigms 2.0 process}
+Like the process that led to Paradigms 1.0, the Paradigms 2.0 process is a ``shared vision'' type of change \cite{henderson2010}.
+Changes to the course structures of the physics curriculum emerged from discussions among the members of the department.
 We began the Paradigms~2.0 process in the Winter of 2016, having in
@@
 integration. Mathematically rigorous derivations are only briefly
-discussed in class, with additional details offered in online video
-format. In summary, we emphasize the physics concepts and reasoning
+mentioned in class. Instead we emphasize the physics concepts and reasoning
 skills that allow professional physicists to quickly/quantitatively
@@
 
-The students engage with the material by solving daily quiz questions
-in class, solving weekly homework problems, performing a modern
-physics experiment, and writing a term paper. The daily quiz questions
-help clarify the goal of each lecture. The weekly homework questions
-are directly related to contemporary challenges requiring a variety
-of estimation techniques. The modern physics experiment, performed
-with a low cost kit (LEDs, diffraction grating, rulers, volt meter
-etc.) emphasizes a quantitative approach to observation and explores
-issues of experimental uncertainty. Lastly, the term paper allows
-students to dive deeper into a particular contemporary challenge. The
-term paper tests students' ability to make their own initial
-assessment of a complex problem, and communicate their ideas with
-diagrams, graphs and estimations.
+% The students engage with the material by solving daily quiz questions
+% in class, solving weekly homework problems, performing a modern
+% physics experiment, and writing a term paper. The daily quiz questions
+% help clarify the goal of each lecture. The weekly homework questions
+% are directly related to contemporary challenges requiring a variety
+% of estimation techniques. The modern physics experiment, performed
+% with a low cost kit (LEDs, diffraction grating, rulers, volt meter
+% etc.) emphasizes a quantitative approach to observation and explores
+% issues of experimental uncertainty. Lastly, the term paper allows
+% students to dive deeper into a particular contemporary challenge. The
+% term paper tests students' ability to make their own initial
+% assessment of a complex problem, and communicate their ideas with
+% diagrams, graphs and estimations.
 
@@
 essence of what makes the Paradigms in Physics program distinctive.
-Emily van Zee, a researcher with experience in ..., 
-\corinnesays{Emily, please add a short description of yourself} 
-has completed a retrospective study of the process we used to develop the 
+One of the authors (EvZ) have completed a retrospective study of the 
+process we used to develop the 
 original Paradigms program and we have used her results extensively in 
@@
 ongoing, as we are having professors new to the department shadow
-experienced faculty as they teach \emph{Paradigm} course, prior to teaching
+experienced faculty as they teach \emph{Paradigm} courses, prior to teaching
 the same course themselves.

### 2017-06-29 paper.tex

@@
 exposes students to increased levels of mathematical sophistication and
-sense-making strategies. \davidsays{ADD CITATIONS SOMEWHERE HERE TO KELBY AND MACKENZIE}
+sense-making strategies. \davidsays{ADD CITATIONS SOMEWHERE HERE TO MACKENZIE}
 
@@
 explicitly in all aspects of the course: exams, homework and during
-in-class activities. \davidsays{CITE HERE?} Our aim is that students will develop
+in-class activities~\cite{hahn2017sensemaking}. Our aim is that students will develop
 sense-making skills that will improve their learning in advanced

### 2017-06-29 paper.tex

@@
 exposes students to increased levels of mathematical sophistication and
-sense-making strategies. \davidsays{ADD CITATIONS SOMEWHERE HERE TO MACKENZIE}
+sense-making strategies.
 
@@
 explicitly in all aspects of the course: exams, homework and during
-in-class activities~\cite{hahn2017sensemaking}. Our aim is that students will develop
+in-class activities~\cite{lenz2017sensemaking, hahn2017sensemaking}. Our aim is that students will develop
 sense-making skills that will improve their learning in advanced

### 2018-09-26 proposal.tex
work on proposal

\documentclass[aps,prstper,reprint]{revtex4-1}   
\usepackage{graphicx}
\usepackage{hyperref}  
\hypersetup{colorlinks=true,urlcolor=blue,citecolor=blue,linkcolor=blue}   
\urlstyle{same}
\pagestyle{empty}		% page numbers added later, when compiling the whole proceedings

\newcommand\corinnesays[1]{\textcolor{red}{[\sc CAM: {\em#1}]}}
\newcommand\davidsays[1]{\textcolor{blue}{[\sc DR: {\em#1}]}}
\renewcommand\davidsays[1]{}

\begin{document}

\title{Paradigms in Physics 2.0}
\author{David Roundy}
\author{Elizabeth Gire}
\author{Ethan Minot}
\author{Emily van Zee}
\affiliation{Department of Physics, Oregon State University, Corvallis, Oregon, 97331}
\author{Corinne A. Manogue}
\affiliation{Department of Physics, Oregon State University, Corvallis, Oregon, 97331}

%\keywords{}

\begin{abstract}
In 2016, the Department of Physics at Oregon State University began a
process to revise our Paradigms in Physics curriculum for physics
majors.  We began with a colloquium to inform the department of our
plans and request their assistance, followed by a
survey of students and faculty as well as individual interviews with
the faculty teaching each course.  As we developed a plan to
address student- and faculty-identified challenges in the curriculum,
we met with each faculty member individually to explain and refine our
proposal, which was unanimously approved by the faculty.  Major
changes include major changes to several courses (math
methods, computational physics, modern physics, electronics, and
classical mechanics), including the introduction of two sophomore-year
courses designed specifically to help prepare students for their
upper-division courses.
\end{abstract}

\maketitle

\section{Theoretical basis for the work}
The framework for this study was adapted from studies of
organizational decision-making\cite{beach2006leadership}. Leaders in
an organization ``must work with others within the organization (1) to
understand the organization's internal and external environments, (2)
to understand the organization's culture, (3) to create a vision of a
desirable future and obtain buy-in, (4) to design a plan that moves
the organization toward the envisioned future, (5) to integrate the
various units in implementation of the plan and to monitor progress
and (6) to institutionalize achieved changes and make continuous
change an integral part of the organization's culture'' (Beach ,
p. ix).  \cite{beach2006leadership}
\section{Research questions}
\begin{itemize}
\item What was the educational reform context within which the faculty
  initiated revising the upper-level physics curriculum?
\item What was the culture of the department at that time? In
  particular, what were the participants' beliefs about teaching and
  learning?
\item What motivated making changes and how did the faculty form a
  coherent vision for the new program?
\item How did the faculty develop and communicate plans for this new
  curriculum?
\item How have they changed instruction based upon the proposals that
  the faculty approved with an unanimous vote?
\end{itemize}
\section{Research methods}
We performed interviews of faculty before, during and after the
Paradigms 2.0 process, gave surveys to students and faculty, and had
an observer take notes during all Paradigms 2.0 meetings.

%% The Paradigms in Physics project began in 1996, when three faculty
%% members at Oregon State applied to the NSF for funding to redesign the
%% upper-division curriculum.  The motivation for the original Paradigms~1.0
%% project was in fact similar to the motivation for Paradigms~2.0: a
%% desire to soften the ``brick wall'' encountered by students upon
%% reaching the junior year, within the constraint that transfer students
%% must be able to graduate with two years of upper-division
%% courses~\cite{manogue2001paradigms}.  On top of this, there was a
%% desire to provide students with a broad knowledge of physics prior to
%% the GRE exams in the fall of their senior year.  The process of
%% redesigning the curriculum was guided by the creation of index cards
%% listing subject content, and sorting these cards into courses.  This
%% process involved the entire faculty, and culminated in a unanimous
%% agreement to adopt the resulting curriculum.

%% The resulting curriculum was primarily composed of intensive junior-year \emph{Paradigm}
%% courses, followed by more conventional senior-year courses.
%% The \emph{Paradigm} courses meet every day for a total of seven
%% hours per week for 3 weeks.  These courses incorporate
%% laboratory experiences and active engagement into the class, and
%% typically have two problem sets per week.  These courses also had
%% integrated math content, and were followed by a \emph{Math Methods} course.
%% The senior-year courses are more traditional 3-credit courses which meet
%% three hours per week for an entire 10-week quarter.

%% Two decades have passed since the original Paradigms~1.0 effort.
%% During this time we have made a number of changes: e.g. a new
%% first \emph{Paradigm} course was introduced to soften the beginning of the
%% junior year, the order of courses was shuffled more than once, \emph{Math
%% Methods} was moved from the beginning of the senior year to the end of
%% the junior year, and a computational laboratory course was introduced
%% to accompany the junior-year \emph{Paradigm} courses.  The faculty maintained the tradition of meeting
%% every three weeks to discuss issues relating to upper-division
%% teaching.  New faculty arrived and learned to teach the new courses,
%% and introduced their own ideas to the courses.

%% %\subsection{Motivation for Paradigms 2.0}
%% A number of factors motivated us to embark on the
%% Paradigms~2.0 process.  In the last two decades, we have observed
%% a number of challenges students face in our major.  While in many cases we
%% addressed these by changing and reordering the courses, we felt a look
%% at the entire curriculum was in order.
%% In addition, our faculty
%% understanding of the details of the sequence was diminishing by attrition: 
%% while our younger
%% faculty are enthusiastic about the Paradigms program, many lack perspective on
%% where students are in a given course, and what content is essential
%% for students to grasp for a subsequent course.

%% \paragraph{Transfer students}
%% Our curriculum is strongly affected by our desire to accommodate
%% transfer students from community colleges.  This leads us to focus on
%% a physics major in which all upper-division courses are taken in two
%% years.  However, our current curriculum is very hard on students in
%% the fall of their junior year, and especially so for transfer
%% students.  So we aimed to structure and order the courses such that non-transfer
%% students could be better prepared for their junior year, while
%% transfer students could postpone a few courses so that the
%% Fall quarter of their junior year would be no more difficult than that of any other student.

%% \paragraph{Coupled curricular changes}
%% A backlog of curricular changes had accumulated that could not be
%% separated from an examination of the curriculum as a whole.  The requirements
%% for computational physics, the number of required electronics credits, and the
%% role of our modern physics course could not easily be addressed separately.

%% \paragraph{New faculty}
%% % Corinne, Henri, David McI, Janet, Tom = 5

%% % David R, Ethan, Oksana, Bo, Guenter, Weihong, Matt, Yun-Shik, Heidi,
%% % Liz, Davide = 11
%% Only 5 of our 16 current tenure-line faculty were involved in the
%% original Paradigms~1.0 effort.  Some of the newest faculty have only a
%% superficial understanding of our course structure.  This poses a
%% challenge when these professors teach courses that are closely
%% intertwined with one another.

%% \section{Paradigms 2.0 process}
%% Like the process that led to Paradigms 1.0, the Paradigms 2.0 process
%% is a ``shared vision'' type of change \cite{henderson2010}. Changes to
%% the course structures of the physics curriculum emerged from
%% discussions among the members of the department. We began the
%% Paradigms~2.0 process in the Winter of 2016, having in the previous
%% year obtained faculty agreement, and support from our Department head.
%% The process was spearheaded by a committee of four (the authors DR,
%% EG, EM, and CAM, with EvZ present at meetings documenting our
%% process).  During the Winter quarter, we informed the community of our
%% process through a colloquium, and collected student and faculty
%% perspectives on the existing curriculum.  We then interviewed faculty
%% who recently taught each of our existing courses to document topics
%% that were currently covered.  This resulted in a total of $\sim$700
%% index cards in $\sim$30 stacks, with each stack corresponding to one
%% course, and each card describing a topic, color-coded as in
%% Fig.~\ref{fig:schedule}.

%% The committee met twice a week to discuss existing challenges and sort
%% the cards into new stacks corresponding to new and reordered courses.
%% When we discussed major changes to a course, we often invited
%% interested faculty to join us to provide their own perspective on
%% possible challenges and improvements.  Once we had a draft proposal,
%% we began inviting each faculty member in to see the cards and discuss
%% the proposed sequence of courses.  In addition, we had a focus group with
%% all the current students to explain the proposed changes and request
%% feedback.

%% After incorporating the feedback from  individual meetings with every
%% faculty member, we scheduled
%% two full faculty meetings with a week in between.  In the first meeting,
%% we presented our proposal in detail and invited questions, but not discussion.
%% This was needed for a couple of reasons:  those faculty we met with first
%% may not have seen the final proposal, and some faculty during their
%% individual meetings chose to focus on a small subset of the
%% curriculum, often relating to courses they had themselves taught.
%% During the following week faculty engaged in hallway discussions of the proposal.
%% This week helped ensure that the faculty did not feel rushed into a vote.
%% In the second faculty meeting, we again presented our proposal, and opened
%% the floor for discussion.  After considerable discussion, largely on
%% changes that were not part of the proposal, the faculty unanimously voted to adopt
%% the proposed changes.

%% \begin{figure}
%% \includegraphics[width=\columnwidth]{vertical-schedule}
%% \caption{Hypothetical student schedule in the new curriculum.
%%   Transfer students begin in year 3, and take the marked additional
%%   courses.\label{fig:schedule}}
%% \end{figure}

\acknowledgments{This work was supported by the National Science
  Foundation under Grant No. 1323800 Supplement.}

% For a longer bibliography, delete the thebibliography block above, then comment in 
% these two lines to use a .bib file with BibTeX.
\bibliographystyle{apsrev}  	% supercedes the longbibliography option, so leave commented out if you want to display article titles
\bibliography{paper}  	% don't include the .bib suffix

\end{document}

### 2018-09-27 proposal.tex
initial proposal

@@
 \urlstyle{same}
-\pagestyle{empty}		% page numbers added later, when compiling the whole proceedings
 
@@
 \newcommand\davidsays[1]{\textcolor{blue}{[\sc DR: {\em#1}]}}
-\renewcommand\davidsays[1]{}
 
@@
 \author{David Roundy}
+\email{roundyd@physics.oregonstate.edu}
 \author{Elizabeth Gire}
@@
 
-\section{Theoretical basis for the work}
-The framework for this study was adapted from studies of
-organizational decision-making\cite{beach2006leadership}. Leaders in
-an organization ``must work with others within the organization (1) to
-understand the organization's internal and external environments, (2)
-to understand the organization's culture, (3) to create a vision of a
-desirable future and obtain buy-in, (4) to design a plan that moves
-the organization toward the envisioned future, (5) to integrate the
-various units in implementation of the plan and to monitor progress
-and (6) to institutionalize achieved changes and make continuous
-change an integral part of the organization's culture'' (Beach ,
-p. ix).  \cite{beach2006leadership}
-\section{Research questions}
+This paper will describe both the process of the reform and its
+outcome, but the emphasis of the paper and the research aspect of it
+will be on the process by which we developed the changes and arrived
+at consensus.  This is paper is thus on the coarse-grained picture of
+how one department acheived a broad-scale reform of its curriculum,
+which we expect will serve as a valuable balance to the more
+fine-grained view that is needed for classroom activities.  In
+addition to the Paradigms 2.0 reform, we will briefly introduce the
+original effort that began in 1996, and give a bit more detail on how
+that reform was maintained and refined during the intervening decades.
+
+This story of how we achieved a dramatic curriculum change will be
+useful for other departments considering a reform of their curriculum.
+Few departments embark on such an endeavor, and publishing a blueprint
+for how we managed this will be helpful for others seeking to reform
+their curriculum.
+
+\paragraph*{Theoretical basis for the work}
+The framework for this {\bf empirical study} was adapted from studies
+of organizational decision-making\cite{beach2006leadership}.  Like the
+process that led to Paradigms 1.0, the Paradigms 2.0 process is a
+``shared vision'' type of change\cite{henderson2010}. Changes to the
+course structures of the physics curriculum emerged from discussions
+among the members of the department. The Paradigms~2.0 process began
+in the Winter of 2016, after the faculty agreed to the process during
+the previous year.
+
+\paragraph*{Research questions}
 \begin{itemize}
@@
 \end{itemize}
-\section{Research methods}
+\paragraph*{Research methods}
 We performed interviews of faculty before, during and after the

### 2018-09-28 proposal.tex
Suggestions from Liz

@@
 
-\begin{abstract}
+\maketitle
+
 In 2016, the Department of Physics at Oregon State University began a
@@
 upper-division courses.
-\end{abstract}
-
-\maketitle
 
@@
 \begin{itemize}
-\item What was the educational reform context within which the faculty
-  initiated revising the upper-level physics curriculum?
-\item What was the culture of the department at that time? In
+\item What was the culture of the department when the faculty initiated revising the upper-level physics curriculum? In
   particular, what were the participants' beliefs about teaching and

### 2018-09-29 proposal.tex
Changes by Emily

@@
 
-\newcommand\corinnesays[1]{\textcolor{red}{[\sc CAM: {\em#1}]}}
+\newcommand\evzsays[1]{\textcolor{red}{[\sc EvZ: {\em#1}]}}
 \newcommand\davidsays[1]{\textcolor{blue}{[\sc DR: {\em#1}]}}
@@
 process to revise our Paradigms in Physics curriculum for physics
-majors.  We began with a colloquium to inform the department of our
-plans and request their assistance, followed by a
-survey of students and faculty as well as individual interviews with
-the faculty teaching each course.  As we developed a plan to
-address student- and faculty-identified challenges in the curriculum,
-we met with each faculty member individually to explain and refine our
-proposal, which was unanimously approved by the faculty.  Major
-changes include major changes to several courses (math
-methods, computational physics, modern physics, electronics, and
-classical mechanics), including the introduction of two sophomore-year
-courses designed specifically to help prepare students for their
-upper-division courses.
+majors.
+Paradigms~2.0 began in the Winter of 2016, after the faculty agreed to undertake this reform during the previous year.
+The process included frequent meetings by a small committee of faculty volunteers, a colloquium to communicate plans and to request assistance, an online survey of students and faculty to identify issues, individual interviews with the faculty teaching each course, and discussions with focus groups of students and instructional faculty.
+
+As we developed ideas to address challenges in the curriculum, we met with each faculty member individually to explain and refine our proposal.
+Major changes included revising a series of courses, eliminating others, re-conceptualizing how to integrate just-in-time mathematics instruction, and introducing two sophomore-year courses designed specifically to help prepare students for their upper-division courses.
+The proposal was unanimously approved by the faculty and then implemented during 2016-2017.
 
 This paper will describe both the process of the reform and its
-outcome, but the emphasis of the paper and the research aspect of it
-will be on the process by which we developed the changes and arrived
-at consensus.  This is paper is thus on the coarse-grained picture of
-how one department acheived a broad-scale reform of its curriculum,
-which we expect will serve as a valuable balance to the more
-fine-grained view that is needed for classroom activities.  In
-addition to the Paradigms 2.0 reform, we will briefly introduce the
-original effort that began in 1996, and give a bit more detail on how
-that reform was maintained and refined during the intervening decades.
-
-This story of how we achieved a dramatic curriculum change will be
-useful for other departments considering a reform of their curriculum.
-Few departments embark on such an endeavor, and publishing a blueprint
-for how we managed this will be helpful for others seeking to reform
-their curriculum.
+outcome.
+It will provide a coarse-grained picture of how one department achieved a broad-scale reform of its upper-division curriculum, which can serve as a valuable balance to the more fine-grained view that is needed for refining classroom activities or revising a single course.
+In addition to the Paradigms 2.0 reform, we will briefly introduce the original effort that began in 1996 \cite{manogue2001paradigms,manogue2003paradigms}, and summarize how that reform was maintained and refined during the intervening decades.
+Few departments embark on such a long-term endeavor and publishing a report on how we managed this may be helpful for others seeking to reform their curriculum.
 
@@
 ``shared vision'' type of change\cite{henderson2010}. Changes to the
-course structures of the physics curriculum emerged from discussions
-among the members of the department. The Paradigms~2.0 process began
-in the Winter of 2016, after the faculty agreed to the process during
-the previous year.
+course structures of the physics curriculum emerged from on-going collaborative discussions among the members of the department.
 
@@
 \item What motivated making changes and how did the faculty form a
-  coherent vision for the new program?
-\item How did the faculty develop and communicate plans for this new
+  coherent vision for the revised program?
+\item How did the faculty develop and communicate plans for this revised
   curriculum?
@@
 \paragraph*{Research methods}
-We performed interviews of faculty before, during and after the
-Paradigms 2.0 process, gave surveys to students and faculty, and had
-an observer take notes during all Paradigms 2.0 meetings.
+Data included audio recordings of interviews with faculty before, during and after the Paradigms 2.0 process; responses to an online survey of students and faculty; audio recordings, notes, and some photos during Paradigms 2.0 Committee meetings, Upper Division Curriculum Committee meetings, and some faculty meetings; and copies of relevant documents.
+Analysis included a quantitative summary of survey data and qualitative interpretations of what happened while initiating and sustaining the Paradigms in Physics program (1996-2015), re-envisioning the Paradigms Program during 2016, and implementing Paradigms 2.0 during 2016-2017 and beyond.
 

### 2018-09-29 proposal.tex
Final changes, from Emily, etc.

@@
 
+\newcommand\camsays[1]{\textcolor{red}{[\sc CAM: {\em#1}]}}
 \newcommand\evzsays[1]{\textcolor{red}{[\sc EvZ: {\em#1}]}}
@@
 In 2016, the Department of Physics at Oregon State University began a
-process to revise our Paradigms in Physics curriculum for physics
+process to revise our award-winning Paradigms in Physics curriculum for physics
 majors.
-Paradigms~2.0 began in the Winter of 2016, after the faculty agreed to undertake this reform during the previous year.
-The process included frequent meetings by a small committee of faculty volunteers, a colloquium to communicate plans and to request assistance, an online survey of students and faculty to identify issues, individual interviews with the faculty teaching each course, and discussions with focus groups of students and instructional faculty.
+Paradigms~2.0 began in the Winter of 2016.
+The process included frequent meetings by a small committee of faculty volunteers, a colloquium to communicate plans and to request assistance, an online survey of students and faculty to identify issues, individual interviews with faculty teaching each course, and discussions with focus groups of students and faculty.
 
-As we developed ideas to address challenges in the curriculum, we met with each faculty member individually to explain and refine our proposal.
+As we developed ideas to address identified challenges in the curriculum, we met with each faculty member individually to explain and refine our proposal.
 Major changes included revising a series of courses, eliminating others, re-conceptualizing how to integrate just-in-time mathematics instruction, and introducing two sophomore-year courses designed specifically to help prepare students for their upper-division courses.
-The proposal was unanimously approved by the faculty and then implemented during 2016-2017.
+The proposal was unanimously approved by the faculty and implemented during 2016-2017.
 
-This paper will describe both the process of the reform and its
+This paper describes both our research about the process of the reform and its
 outcome.
-It will provide a coarse-grained picture of how one department achieved a broad-scale reform of its upper-division curriculum, which can serve as a valuable balance to the more fine-grained view that is needed for refining classroom activities or revising a single course.
-In addition to the Paradigms 2.0 reform, we will briefly introduce the original effort that began in 1996 \cite{manogue2001paradigms,manogue2003paradigms}, and summarize how that reform was maintained and refined during the intervening decades.
-Few departments embark on such a long-term endeavor and publishing a report on how we managed this may be helpful for others seeking to reform their curriculum.
+It shows how one department achieved a broad-scale reform of its upper-division curriculum, which can serve as a valuable balance to the more fine-grained view that is needed for designing/refining classroom activities or revising a single course.
+In addition, we briefly summarize how the original effort began\cite{manogue2001paradigms,manogue2003paradigms}, was initiated and sustained over  two decades\cite{vanzeeHistory}.
+Few departments embark on such a long-term endeavor and publishing a report on how we managed this may be helpful for others seeking to reform their curriculum and sustain their changes in a flexible, organic way.
 
@@
 \begin{itemize}
-\item What was the culture of the department when the faculty initiated revising the upper-level physics curriculum? In
-  particular, what were the participants' beliefs about teaching and
-  learning?
+\item What was the culture of the department when the faculty initiated revising the upper-level physics curriculum? In particular, what were the participants' beliefs about teaching and learning?
 \item What motivated making changes and how did the faculty form a
@@
 \item How have they changed instruction based upon the proposals that
-  the faculty approved with an unanimous vote?
+  the faculty approved with a unanimous vote?
 \end{itemize}
@@
 
-\acknowledgments{This work was supported by the National Science
-  Foundation under Grant No. 1323800 Supplement.}
+\acknowledgments{This work was supported by the National Science Foundation under Grant No. 1323800 Supplement.}
 
